---
output: 
  html_document:
      toc: false
      toc_float: false
      toc_depth: 4
      code_folding: NULL
      df_print: paged
editor_options: 
  chunk_output_type: console
---

<!-- This script reads from a folder containing (VEP+snpEff) annotated CAW (aka Sarek) output, including copy number from ControlFreec and Ascat, structural variation from Manta, somatic point mutations from Mutact 2 and Strelka, and germline mutations from HaplotypeCaller.-->

<!-- CSS style for left alignment and for tables to look ok: -->
<style>
body {
position: absolute;
left: 42px;}
.main-container {max-width: none !important;}
table, td, th {
font-size: 11px;
border: none;
padding-left: 1em;
padding-right: 1em;
min-width: 100%;
word-wrap: break-word;
max-width: 240px;
margin-left: auto;
margin-right: auto;
margin-top: 1em;
margin-bottom: 1em;
}

</style>

<!-- Markdown setup -->
```{r setup, include=F}
knitr::opts_chunk$set(echo = FALSE,comment = '',warning=F,error = F)
options(width = 600)
```

# {.tabset .tabset-pills}

```{r project_and_refgenome}
cat('Sample:',basename(getwd()),'\tDirectory:',getwd(),'\nDate:',date())

## Project and reference genome

project='BTB'
#project='AML-p'
#project='AML-a'
#project='CLL'

if (project=='BTB') reference_genome='GRCh38' else reference_genome='hg19'

if (project=='BTB' | project=='CLL') write_tables=T else write_tables=F

```

## Setup {.tabset}

### List files {.active}
<!-- First, make a list of files to read and visualize -->
```{r list_files}


if (project == 'CLL') {
  name=basename(getwd())
  path=paste0('/castor/project/proj/WGS2018Work/WGS_Subsets/',name,'/results')
  # Ascat files
  ascat_Tratio_file <- dir(path=path,pattern=paste0('^T.*LogR$'),recursive = T,full.names = T)
  ascat_Tbaf_file <- dir(path=path,pattern=paste0('^T.*BAF$'),recursive = T,full.names = T)
  ascat_Nbaf_file <- dir(path=path,pattern=paste0('^N.*BAF$'),recursive = T,full.names = T)
  ascat_segment_file <- dir(path=path,pattern=paste0('^T.*cnvs.txt$'),recursive = T,full.names = T)
  # Manta structural variant files
  manta_tumor_file <- dir(path=path,pattern='.*somaticSV.vcf.snpEff.ann.vcf$',recursive = T,full.names = T)
  manta_normal_file <- dir(path=path,pattern='.*diploidSV.vcf.snpEff.ann.vcf$',recursive = T,full.names = T)[1]

  # VEP Strelka snv file
  strelka_snv_file <- dir(path=path,pattern=paste0('Strelka.*somatic_snvs.*vep.ann.vcf$'),recursive = T,full.names = T)
  # VEP Strelka indel file
  strelka_indel_file <- dir(path=path,pattern=paste0('Strelka.*somatic_indels.*vep.ann.vcf$'),recursive = T,full.names = T)
}

if (project == 'AML-a') {
  name=basename(getwd())
  # Ascat files
  ascat_Tratio_file <- dir(path='~/results/VariantCalling/Ascat/',pattern=paste0('.*',name,'-[DR].*LogR$'),recursive = T,full.names = T)
  ascat_Tbaf_file <- dir(path='~/results/VariantCalling/Ascat/',pattern=paste0('.*',name,'-[DR].*BAF$'),recursive = T,full.names = T)
  ascat_Nbaf_file <- dir(path='~/results/VariantCalling/Ascat/',pattern=paste0('.*',name,'-G.BAF$'),recursive = T,full.names = T)
  ascat_segment_file <- dir(path='~/results/VariantCalling/Ascat/',pattern=paste0('.*',name,'-[DR].*cnvs.txt$'),recursive = T,full.names = T)
  # Manta structural variant files
  manta_tumor_file <- dir(path='~/results/Annotation/SnpEff',
                          pattern=paste0('Manta.*',name,'.*[DR].*somaticSV.vcf.snpEff.ann.vcf$'),recursive = T,full.names = T)
  manta_normal_file <- dir(path='~/results/Annotation/SnpEff',
                          pattern=paste0('Manta.*',name,'.*diploidSV.vcf.snpEff.ann.vcf$'),recursive = T,full.names = T)[1]
  # Haplotype caller variant files
  #haplotypecaller_T_file <- grep(pattern = ".*-G.clean.dedup.recal.bam.raw.annotated.vcf.vep.ann.vcf$",files,value = T)
  #haplotypecaller_N_file <- dir(path='~/results/Annotation/VEP',pattern=paste0('.*',name,'.*vep.ann.vcf$'),recursive = T,full.names = T)[1]
  # Strelka variant files
  # VEP Strelka snv file
  strelka_snv_file <- dir(path='~/results/Annotation/VEP',
                          pattern=paste0('Strelka.*',name,'.*[DR].*.somatic_snvs.vcf.vep.ann.vcf$'),recursive = T,full.names = T)
  # VEP Strelka indel file
  strelka_indel_file <- dir(path='~/results/Annotation/VEP',
                          pattern=paste0('Strelka.*',name,'.*[DR].*.somatic_indels.vcf.vep.ann.vcf$'),recursive = T,full.names = T)
}


if (project == 'AML-p') {
  files <- dir(recursive = T,full.names = T)

  files <- files[-grep(pattern = '.png',x = files)]
  files <- files[-grep(pattern = '.gz',x = files)]  
  
  # Ascat files
  ascat_Tratio_file <- files[grep(pattern = "^.*Ascat.*-[DR].*LogR$",files)]
  ascat_Tbaf_file <- files[grep(pattern = "^.*Ascat.*-[DR].*BAF$",files)]
  ascat_Nbaf_file <- files[grep(pattern = "^.*Ascat.*-G.*BAF$",files)][1]
  ascat_segment_file <- files[grep(pattern = "^.*Ascat.*-[DR].*\\.cnvs\\.txt$",files)]
  # Manta structural variant files
  manta_tumor_file <- grep(pattern = ".*Manta_.*vs.*somaticSV.vcf.snpEff.ann.vcf$",files,value = T)
  manta_normal_file <- grep(pattern = ".*Manta_.*vs.*diploidSV.vcf.snpEff.ann.vcf$",files,value = T)[1]
  # Haplotype caller variant files
  #haplotypecaller_T_file <- grep(pattern = ".*-G.clean.dedup.recal.bam.raw.annotated.vcf.vep.ann.vcf$",files,value = T)
  haplotypecaller_N_file <- grep(pattern = ".*-G.clean.dedup.recal.bam.raw.annotated.vcf.vep.ann.vcf$",files,value = T)
  # Strelka variant files
  # snpEff + VEP Strelka snv file
  strelka_snv_file <- grep(pattern = ".*Strelka_.*_somatic_snvs.*vep.ann.vcf$",files,value = T)
  # snpEff + VEP Strelka indel file
  strelka_indel_file <- grep(pattern = ".*Strelka_.*_somatic_indels.*vep.ann.vcf$",files,value = T)
}

if (project == 'BTB') {
  files <- dir(recursive = T,full.names = T)

  remove=unique(c(grep(pattern = '.png',x = files),
                  #grep(pattern = 'ControlFreec',x = files),
                  grep(pattern = 'Annotation.old',x = files),
                  grep(pattern = '/work/',x = files)))
  
  if (length(remove)>0) files <- files[-remove]
  
  # control freec files
  freec_Tratio_file <- files[grep(pattern = '[TR].hg38.pileup.gz_ratio.txt',files,perl = T)]
  freec_Nratio_file <- files[grep(pattern = '[TR].hg38.pileup.gz_normal_ratio.txt',files,perl = T)][1]
  freec_Tbaf_file <- files[grep(pattern = '[TR].hg38.pileup.gz_BAF.txt',files,perl = T)]
  freec_Nbaf_file <- files[grep(pattern = '[BN].hg38.pileup.gz_BAF.txt',files,perl = T)][1]
  freec_info_file <- files[grep(pattern = 'hg38.pileup.gz_info.txt',files,perl = T)]
  freec_cnv_file <- files[grep(pattern = "^.*[TR]\\.hg38\\.pileup\\.gz_CNVs$",files,perl = T)]
  # Ascat files
  ascat_Tratio_file <- files[grep(pattern = "^.*Ascat.*[TR]\\.LogR$",files)]
  ascat_Nratio_file <- files[grep(pattern = "^.*Ascat.*[BN]\\.LogR$",files)][1]
  ascat_Tbaf_file <- files[grep(pattern = "^.*Ascat.*[TR]\\.BAF$",files)]
  ascat_Nbaf_file <- files[grep(pattern = "^.*Ascat.*[BN]\\.BAF$",files)][1]
  ascat_segment_file <- files[grep(pattern = "^.*Ascat.*[TR]\\.cnvs\\.txt$",files)]
  # ascat_Tratio_file <- files[grep(pattern = "^.*Ascat.*02\\.LogR$",files)]
  # ascat_Nratio_file <- files[grep(pattern = "^.*Ascat.*01\\.LogR$",files)]
  # ascat_Tbaf_file <- files[grep(pattern = "^.*Ascat.*02\\.BAF$",files)]
  # ascat_Nbaf_file <- files[grep(pattern = "^.*Ascat.*01\\.BAF$",files)]
  # ascat_segment_file <- files[grep(pattern = "^.*Ascat.*02\\.cnvs\\.txt$",files)]
  # Manta structural variant files
  manta_tumor_file <- grep(pattern = ".*Manta_.*vs.*somaticSV.vcf.snpEff.ann.vcf$",files,value = T)
  manta_normal_file <- grep(pattern = ".*Manta_.*vs.*diploidSV.vcf.snpEff.ann.vcf$",files,value = T)[1]
  # Haplotype caller variant files
  haplotypecaller_T_file <- grep(pattern = ".*haplotypecaller.*[TR].*AF.*vep.ann.vcf$",files,value = T)
  # haplotypecaller_T_file <- 
  #   grep(pattern = "^.*Annotation/VEP/haplotypecaller.*02.*vcf$",files,value = T)[1]
  haplotypecaller_N_file <- grep(pattern = ".*haplotypecaller.*[NB].*AF.*vep.ann.vcf$",files,value = T)
  # haplotypecaller_N_file <- 
  #   grep(pattern = "^.*Annotation/VEP/haplotypecaller.*01.*vcf$",files,value = T)[1]
  # snpEff+VEP Mutect2 file
  # mutect2_file <- grep(pattern = "^.*Annotation/VEP/mutect2_.*[TR]_vs_.*[BN]\\.vcf.*vcf$",files,value = T)[1]
  mutect2_file <- grep(pattern = ".*mutect2_.*AF.*vep.ann.vcf$",files,value = T)
  # snpEff + VEP Strelka snv file
  strelka_snv_file <- grep(pattern = ".*Strelka_.*_somatic_snvs.*AF.*vep.ann.vcf$",files,value = T)
  # snpEff + VEP Strelka indel file
  strelka_indel_file <- grep(pattern = ".*Strelka_.*_somatic_indels.*AF.*vep.ann.vcf$",files,value = T)
}


if (exists('files')) rm(files)
{
  cat('Selected files in',getwd(),'\n\n')
  for ( obj in ls() ) { 
    cat('---',obj,'---\n') 
    print(get(obj)) 
  }
}  


```


### Load dependencies
<!-- Dependencies -->
```{r load_dependencies}
library(data.table)
library(ggplot2); theme_set(theme_light())
library(grid)
library(dplyr)
library(ggrepel)
library(VariantAnnotation)
library(VennDiagram)
library(knitr)
library(htmlTable)
library(kableExtra)
library(stringr)

# this is in preparation for VCF file output:
# info_header_add=data.frame(Number=1,
#                            Type=c('Integer','String','Integer','String'),
#                            Description=c('Swegen observations','Concomitant loss of heterozygosity','Pathfindr rank score', 'Pathfindr rank terms'),
#                            row.names = c('Swegen_count','LOH','Rank_score','Rank_terms'))


makelinks <- function(strings,sep='&') {
  if (length(strings)>0) for (i in 1:length(strings)) if (strings[i]!='') {
    t=strsplit(strings[i],sep)[[1]]
    for (j in 1:length(t)) {
      if (t[j] %in% alltumorgenes) {
        t[j]=cell_spec(t[j], link = paste0("https://cancer.sanger.ac.uk/cosmic/gene/analysis?ln=",t[j]),format = 'html')
      } else if (startsWith(t[j],'rs')) {
        t[j]=cell_spec(t[j], link = paste0("https://www.ncbi.nlm.nih.gov/projects/SNP/snp_ref.cgi?rs=",substr(t[j],3,100)),format = 'html')
      } else if (startsWith(t[j],'COSM')) {
        t[j]=cell_spec(t[j], link = paste0("https://cancer.sanger.ac.uk/cosmic/mutation/overview?id=",substr(t[j],5,100)),format = 'html')
      }
    }
    strings[i]=paste(t,collapse = ' ')
  }
  return(strings)
}

tableWrapper <- function(table) {
  if ('Existing_variation' %in% colnames(table)) {
    table$Existing_variation=makelinks(table$Existing_variation)
  }
  if ('Gene_Name' %in% colnames(table)) {
    table$Gene_Name[nchar(table$Gene_Name)>200]='many'
  }
  if ('SYMBOL' %in% colnames(table)) {
    table$SYMBOL=makelinks(table$SYMBOL)
  }
  if ('cancer_genes' %in% colnames(table)) {
    table$cancer_genes=makelinks(table$cancer_genes,sep=' ')
  }
  if ('AD_TUMOR' %in% colnames(table)) {
    table$AD_TUMOR=unlist(lapply(table$AD_TUMOR,paste,collapse=', '))
  }
  if ('AD_NORMAL' %in% colnames(table)) {
    table$AD_NORMAL=unlist(lapply(table$AD_NORMAL,paste,collapse=', '))
  }
  if ('AD' %in% colnames(table)) {
    if (is.list(table$AD)) table$AD=unlist(lapply(table$AD,paste,collapse=', '))
  }
  
  samples=sort(unique(table$sample))
  
  if (T) htmlTable(table, 
            col.rgroup = c("none", "#F7F7F7"),
            tspanner=paste('Rank score:',unique(table$rank_score)),
            n.tspanner=rev(table(table$rank_score)))
  # if (length(samples)>1) htmlTable(table, 
  #           col.rgroup = c("none", "#F7F7F7"),
  #           n.rgroup='',
  #           tspanner=paste('Rank score:',unique(table$rank_score)),
  #           n.tspanner=table(table$rank_score))
}

sampleData <- data.table(
  date=date(),
  directory=getwd(),
  name=basename(getwd())#,
  # patient='',
  # sample='',
  # freec_cnv_file,
  # freec_Tratio_file,
  # freec_Nratio_file,
  # freec_Tbaf_file,
  # ascat_Tratio_file
)
if (reference_genome=='GRCh38') chrsz <- data.table(
  chr = paste0('chr',c("1", "2", "3", "4", "5", "6", "7", "8", 
          "9", "10", "11", "12", "13", "14", "15", "16", "17", "18", "19", 
          "20", "21", "22", "X", "Y")), 
  label = c("1", "2", "3", "4", "5", 
          "6", "7", "8", "9", "10", "11", "12", "13", 
          "14", "15", "16", "17", "18", "19", "20", 
          "21", "22", "X", "Y"),
  starts = c(0, 253902921, 500669562, 
             703689723, 898025604, 1084280925, 1259870526, 1424144247, 1574191848, 
             1717288929, 1855896810, 1995944331, 2134165212, 2253485013, 2363996694, 
             2470839615, 2565902256, 2654091777, 2739309138, 2802869979, 2871941700, 
             2923598421, 2979326382, 3140008743), 
  length = c(248902921, 241766641, 
             198020161, 189335881, 181255321, 170589601, 159273721, 145047601, 
             138097081, 133607881, 135047521, 133220881, 114319801, 105511681, 
             101842921, 90062641, 83189521, 80217361, 58560841, 64071721, 
             46656721, 50727961, 155682361, 56827081)
) else chrsz <- data.table(
  chr = c("1", "2", "3", "4", "5", 
          "6", "7", "8", "9", "10", "11", "12", "13", 
          "14", "15", "16", "17", "18", "19", "20", 
          "21", "22", "X", "Y"), 
  label = c("1", "2", "3", "4", "5", 
          "6", "7", "8", "9", "10", "11", "12", "13", 
          "14", "15", "16", "17", "18", "19", "20", 
          "21", "22", "X", "Y"),
  starts = c(0, 249250621, 492449994, 
             690472424, 881626700, 1062541960, 1233657027, 1392795690, 1539159712, 
             1680373143, 1815907890, 1950914406, 2084766301, 2199936179, 2307285719, 
             2409817111, 2500171864, 2581367074, 2659444322, 2718573305, 2781598825, 
             2829728720, 2881033286, 3036303846), 
  length = c(249250621, 243199373, 
             198022430, 191154276, 180915260, 171115067, 159138663, 146364022, 
             141213431, 135534747, 135006516, 133851895, 115169878, 107349540, 
             102531392, 90354753, 81195210, 78077248, 59128983, 63025520, 
             48129895, 51304566, 155270560, 59373566)
  )

```

<!-- Reference data files -->
```{r load_reference}
snptable=fread('~/reports/reference_data/swegen_snp_counts.csv',key='name')
swegen_manta_all=fread('~/reports/reference_data/swegen_sv_counts.csv',key='name')
#allgenes=fread('~/reports/reference_data/genes_ensemble_canonical.csv',key='name')
tumorgenes=fread('~/reports/reference_data/cancer_gene_census.csv',key='Gene Symbol')
tumorgenes$chr=paste0('chr',str_replace(string=tumorgenes$`Genome Location`,pattern = ':.*',replacement = ''))
temp=str_replace(string=tumorgenes$`Genome Location`,pattern = '.*:',replacement = '')
tumorgenes$start=as.numeric(str_replace(string=temp,pattern = '.*-',replacement = ''))
tumorgenes$end=as.numeric(str_replace(string=temp,pattern = '-.*',replacement = ''))
tumorgenes$cumstart=NA; tumorgenes$cumend=NA
for (i in 1:nrow(chrsz)) {
    ix <- tumorgenes$chr==chrsz$chr[i]
    tumorgenes$cumstart[ix] <- tumorgenes$start[ix]+chrsz$starts[i]
    tumorgenes$cumend[ix] <- tumorgenes$end[ix]+chrsz$starts[i]
  }

if (project=='BTB') {
  local_tumorgenes=fread('~/reports/reference_data/2018_gene_list_tere_ref.csv',key='Gene')[Gene!='',1:2]
  alltumorgenes=unique(c(tumorgenes$`Gene Symbol`,local_tumorgenes$Gene))
  alltier1=union(tumorgenes[Tier==1,`Gene Symbol`],local_tumorgenes[`Tier 1 and 2 for pediatric cancers final`==1,Gene])
  alltier2=union(tumorgenes[Tier==2,`Gene Symbol`],local_tumorgenes[`Tier 1 and 2 for pediatric cancers final`==2,Gene])
} else {
  alltumorgenes=unique(c(tumorgenes$`Gene Symbol`))
  alltier1=tumorgenes[Tier==1,`Gene Symbol`]
  alltier2=tumorgenes[Tier==2,`Gene Symbol`]
}

allfusion=tumorgenes[grep('fusion',`Role in Cancer`),`Gene Symbol`]
allfusionpairs=NULL
for (i in 1:length(allfusion)) {
  t=trimws(strsplit(tumorgenes[allfusion[i],`Translocation Partner`],', ')[[1]])
  if (length(t)>0) for (j in 1:length(t))
  allfusionpairs=c(allfusionpairs,paste(sort(c(allfusion[i],t[j])),collapse = ' '))
}
alltsg=tumorgenes[grep('TSG',`Role in Cancer`),`Gene Symbol`]
hotspots_inframe=unique(fread('~/reports/reference_data/hotspots_v2_inframe.csv')[,.(Hugo_Symbol,Amino_Acid_Position)])
hotspots_inframe$start=as.numeric(str_replace(string = hotspots_inframe$Amino_Acid_Position,pattern = '-[0-9]*',replacement = ''))
hotspots_inframe$end=as.numeric(str_replace(string = hotspots_inframe$Amino_Acid_Position,pattern = '[0-9]*-',replacement = ''))
hotspots_snv=unique(
  fread('~/reports/reference_data/hotspots_v2_snv.csv')[,.(Hugo_Symbol,Amino_Acid_Position)])[-grep('splice',Amino_Acid_Position)]
hotspots_snv$pos <- as.numeric(hotspots_snv$Amino_Acid_Position)
near_hotspots=NULL
for (i in 1:nrow(hotspots_inframe)) 
  near_hotspots=c(near_hotspots,
                      paste(hotspots_inframe$Hugo_Symbol[i],
                                 c(seq(hotspots_inframe$start[i]-2,hotspots_inframe$start[i]+2),
                                       hotspots_inframe$end[i]-2,hotspots_inframe$end[i]+2)))
for (i in 1:nrow(hotspots_snv))
  near_hotspots=c(near_hotspots,
                      paste(hotspots_snv$Hugo_Symbol[i],
                                 seq(hotspots_snv$pos[i]-2,hotspots_snv$pos[i]+2)))
cosmic_coding=fread('~/reports/reference_data/cosmic_coding_table.csv',key = 'name')
cosmic_noncoding=fread('~/reports/reference_data/cosmic_noncoding_table.csv',key = 'name')
cosmic_fusions=fread('~/reports/reference_data/cosmic_fusions_table.csv',key = 'name')
```


## Copy number {.tabset}

### ControlFreec
<!-- Read Control Freec copy number data -->
```{r load_control_freec}
cnvs=NULL
freec_cnv=NULL
samplename=NULL
if (project=='BTB') if (!is.na(freec_Tratio_file[1])) {
  # extract sample names
  for (s in 1:length(freec_Tratio_file)) samplename[s]=strsplit(basename(freec_Tratio_file[s]),'[.]')[[1]][1]
  
  tratio=NULL; tbaf=NULL
  for (s in 1:length(freec_Tratio_file)) {
    # tumor log ratio
    temp <- fread(file = freec_Tratio_file[s])
    temp$sample=samplename[s]
    temp=temp[,c(10,1:6)]
    temp$CopyNumber[1:2] <- c(0,4) # to secure range of colors
    tratio=rbind(tratio,temp)
    setkey(tratio,'sample')
    
    # tumor BAF
    temp <- fread(freec_Tbaf_file[s])[,1:3]
    temp$sample=samplename[s]
    temp=temp[,c(4,1:3)]
    tbaf=rbind(tbaf,temp)
    setkey(tbaf,'sample')
  }
  # normal log ratio
  nratio <- fread(file = freec_Nratio_file)[,1:6]
  # normal BAF
  nbaf <- fread(freec_Nbaf_file)[,1:3]

  # manipulate for plot:
  tratio$cumstart <- tratio$Start
  nratio$cumstart <- nratio$Start
  tbaf$cumstart <- tbaf$Position
  nbaf$cumstart <- nbaf$Position
  tempchr=substr(chrsz$chr,4,6)
  for (i in 2:nrow(chrsz)) {
    ix <- tratio$Chromosome==tempchr[i]
    tratio$cumstart[ix] <- tratio$Start[ix]+chrsz$starts[i]
    ix <- nratio$Chromosome==tempchr[i]
    nratio$cumstart[ix] <- nratio$Start[ix]+chrsz$starts[i]
    ix <- tbaf$Chromosome==tempchr[i]
    tbaf$cumstart[ix] <- tbaf$Position[ix]+chrsz$starts[i]
    ix <- nbaf$Chromosome==tempchr[i]
    nbaf$cumstart[ix] <- nbaf$Position[ix]+chrsz$starts[i]
  }
  
  # modify for plot
  tratio$CopyNumber[tratio$CopyNumber>4] <- 4
  nratio$CopyNumber[nratio$CopyNumber>4] <- 4
  tratio$BAF[tratio$BAF<0.5 | tratio$BAF>1] <- NA
  nratio$BAF[nratio$BAF<0.5 | nratio$BAF>1] <- NA
  tratio$Chromosome=paste0('chr',tratio$Chromosome)
  nratio$Chromosome=paste0('chr',nratio$Chromosome)
  tbaf$Chromosome=paste0('chr',tbaf$Chromosome)
  nbaf$Chromosome=paste0('chr',nbaf$Chromosome)
  
  # smooth data for plot
  binned <- NULL
  for (sample in samplename) for (i in 1:nrow(chrsz)) {
    temp <- data.table(
      sample,
      chr=chrsz$chr[i],
      pos=seq(5e5,chrsz$length[i],5e5),
      cumpos=seq(5e5,chrsz$length[i],5e5)+chrsz$starts[i],
      tratio=NA,
      tmaf=NA)
    ctratio <- tratio[sample][Chromosome==chrsz$chr[i]]
    for (j in 1:nrow(temp)) {
      ix <- ctratio$Start>temp$pos[j]-5e5 & ctratio$Start<temp$pos[j]+5e5
      if (sum(ix)>20) {
        d <- density(ctratio$Ratio[ix],na.rm=T)
        temp$tratio[j] <- d$x[which.max(d$y)]
        t=ctratio$BAF[ix]
        t=t[!is.na(t)]
        if (length(t)>10) {
          d <- density(ctratio$BAF[ix],na.rm=T)
          temp$tmaf[j] <- d$x[which.max(d$y)]
        }
      }
    }
    binned <- rbind(binned,temp)
  }
  setkey(binned,'sample')
  

  # check which regions are worth reporting
  # cnv file
  cnvs=NULL
  for (s in 1:length(samplename)) {
    temp <- as.data.table(read.csv(freec_cnv_file[s],sep='\t',header=F))[,1:8]
    names(temp) <- c('chr','start','end','copies','effect','genotype','value','type')
    temp$sample=samplename[s]
    temp=temp[,c(9,1:8)]
    cnvs=rbind(cnvs,temp)
    setkey(cnvs,'sample')
  }
  freec_loh=cnvs[grep('^A*$',genotype)][type=='somatic'][copies<4]
  freec_loh$chr=paste0('chr',freec_loh$chr)
  cnvs=cnvs[effect!='neutral'] # this removes LOH from CNV table
  cnvs$length=paste0(round((cnvs$end-cnvs$start)/1e6,2),'M')
  cnvs$rank_score <- 0
  cnvs$rank_terms <- ''
  cnvs$cancer_genes <- ''
  
  for (i in cnvs[end-start < 50e6, .I]) {
    genes=tumorgenes[chr==paste0('chr',cnvs$chr[i]) & start<cnvs$end[i] & end>cnvs$start[i],`Gene Symbol`] # genes in segment
    if (length(genes)>0) { # if any of our cancer genes
      cnvs$cancer_genes[i] <- paste(genes,collapse = ' ') # put in table
      if (any(genes %in% alltier1,na.rm = T)) {
        cnvs$rank_score[i]=2
        cnvs$rank_terms[i]='T1_gene'
      } else if (any(genes %in% alltier2,na.rm = T)) {
        cnvs$rank_score[i]=2
        cnvs$rank_terms[i]='T2_gene'
      }
    }
    
    # if the effect is focal
    if (cnvs[i,end-start]<3e6) {
      cnvs$rank_score[i]=cnvs$rank_score[i]+1
      cnvs$rank_terms[i]=paste(cnvs$rank_terms[i],'focal')
    }
    
    # if the effect is loss/gain/amp etc
    if (cnvs$copies[i]==0) {
      cnvs$rank_score[i]=cnvs$rank_score[i]+2
      cnvs$rank_terms[i]=paste(cnvs$rank_terms[i],'hz_loss')
    } else if (cnvs$copies[i] > 5) {
      cnvs$rank_score[i]=cnvs$rank_score[i]+2
      cnvs$rank_terms[i]=paste(cnvs$rank_terms[i],'high_amp')
    } else if (cnvs$copies[i] > 2) {
      cnvs$rank_score[i]=cnvs$rank_score[i]+1
      cnvs$rank_terms[i]=paste(cnvs$rank_terms[i],'dup')
    } else if (cnvs$effect[i] =='loss') {
      cnvs$rank_score[i]=cnvs$rank_score[i]+1
      cnvs$rank_terms[i]=paste(cnvs$rank_terms[i],'del')
    }
  }
  
  freec_cnv=cnvs[order(rank_score,decreasing = T)] # order by rank
  if (write_tables) fwrite(freec_cnv,file=paste0(sampleData$name,'_freec_cnv.csv')) # write to file
  
  tableWrapper(freec_cnv) # table in report
}
```

### Ascat
<!-- Read ASCAT copy number data -->
```{r load_ascat}

ascat_cnv=NULL
samplename=NULL
if (length(ascat_Tratio_file)>0) {
  # extract sample names
  for (s in 1:length(ascat_Tratio_file)) samplename[s]=strsplit(basename(ascat_Tratio_file[s]),'[.]')[[1]][1]
  
  ascat_tratio=NULL; ascat_tbaf=NULL; ascat_cnv=NULL
  for (s in 1:length(samplename)) {
    # segmented copy number
    temp=fread(file = ascat_segment_file[s])
    temp$sample=samplename[s]
    temp=temp[,c(6,1:5)]
    temp$chr=paste0('chr',temp$chr)
    ascat_cnv=rbind(ascat_cnv,temp)
    setkey(ascat_cnv,'sample')
    these_cnvs=temp # for use below
    # tumor log ratio
    temp=fread(file = ascat_Tratio_file[s])[,-1]
    #if (project=='AML') temp$Chr=paste0('chr',temp) 
    colnames(temp)[3]='LogR'
    temp$sample=samplename[s]
    temp=temp[order(Chr,Position),c(4,1:3)]
    temp$Ratio=2^temp$LogR
    temp$smoothed=runmed(x = temp$Ratio,k = 99)
    temp$CopyNumber=2
    temp$MinorCopy=1
    for (i in 1:nrow(these_cnvs)) {
      ix=temp$Chr==these_cnvs$chr[i] & temp$Position > these_cnvs$start[i] & temp$Position < these_cnvs$end[i]
      if (!any(ix)) 
        ix=temp$Chr==substr(these_cnvs$chr[i],4,6) & temp$Position > these_cnvs$start[i] & temp$Position < these_cnvs$end[i]
      temp$CopyNumber[ix]=these_cnvs$nMajor[i]+these_cnvs$nMinor[i]
      temp$MinorCopy[ix]=these_cnvs$nMinor[i]
    }
    temp$CopyNumber[temp$CopyNumber>4]=4
    temp$CopyNumber[1:2]=c(0,4)
    ascat_tratio=rbind(ascat_tratio,temp)
    setkey(ascat_tratio,'sample')
    # tumor BAF
    temp=fread(ascat_Tbaf_file[s])[,-1]
    colnames(temp)[3]='BAF'
    temp=temp[which(BAF!=0 & BAF!=1)]
    temp$sample=samplename[s]
    temp=temp[order(Chr,Position),c(4,1:3)]
    ascat_tbaf=rbind(ascat_tbaf,temp)
    setkey(ascat_tbaf,'sample')
  }
  # normal BAF
  ascat_nbaf=fread(ascat_Nbaf_file)[,-1]
  colnames(ascat_nbaf)[3]='BAF'
  ascat_nbaf=ascat_nbaf[which(BAF!=0 & BAF!=1)]

  # join (T) log ratio and (T) BAF
  #ascat_tratio=merge(ascat_tratio,ascat_tbaf,all.x=T)
  #ascat_tratio=merge(ascat_tratio,ascat_nbaf,all.x=T)
  
  # manipulate for plot:
  ascat_tratio$cumstart=ascat_tratio$Position
  ascat_tbaf$cumstart=ascat_tbaf$Position
  ascat_nbaf$cumstart=ascat_nbaf$Position
  #ascat_cnv$LOH=''; ascat_cnv$LOH[ascat_cnv$nMinor==0]='LOH' <---- will use ascat_LOH instead
  for (i in 2:nrow(chrsz)) {
    ix=ascat_tratio$Chr==chrsz$chr[i]
    ascat_tratio$cumstart[ix]=ascat_tratio$Position[ix]+chrsz$starts[i]
    #no n...
    ix=ascat_tbaf$Chr==chrsz$chr[i]
    ascat_tbaf$cumstart[ix]=ascat_tbaf$Position[ix]+chrsz$starts[i]
    ix=ascat_nbaf$Chr==chrsz$chr[i]
    ascat_nbaf$cumstart[ix]=ascat_nbaf$Position[ix]+chrsz$starts[i]
  }
 

  
  # smooth data for view
  ascat_binned=NULL
  for (sample in samplename) for (i in 1:nrow(chrsz)) {
    temp=data.table(
      sample,
      chr=chrsz$chr[i],
      pos=seq(5e5,chrsz$length[i],5e5),
      cumpos=seq(5e5,chrsz$length[i],5e5)+chrsz$starts[i],
      tratio=NA,
      tmaf=NA)
    
    tempLogR=ascat_tratio[sample][Chr==chrsz$chr[i]]
    tempBAF=ascat_tbaf[sample][Chr==chrsz$chr[i]]
    tempBAF$BAF=0.5+abs(tempBAF$BAF-0.5)
    for (j in 1:nrow(temp)) {
      ix = tempLogR$Position>temp$pos[j]-5e5 & tempLogR$Position<temp$pos[j]+5e5
      if (sum(ix)>10) {
        d <- density(2^tempLogR$LogR[ix],na.rm=T)
        temp$tratio[j] <- d$x[which.max(d$y)]
      }
      ix = tempBAF$Position>temp$pos[j]-5e5 & tempBAF$Position<temp$pos[j]+5e5
      if (sum(!is.na(tempBAF$BAF[ix]))>20) {
        d=density(tempBAF$BAF[ix],na.rm=T)
        temp$tmaf[j]=d$x[which.max(d$y)]
      }
    }
    ascat_binned=rbind(ascat_binned,temp)
  }
  setkey(ascat_binned,'sample')

  
  ascat_loh=ascat_cnv[nMinor==0 & nMajor<4]
  ascat_loh$chr=paste0('chr',ascat_loh$chr)
  if (!exists('freec_loh')) freec_loh=ascat_loh # to make sure LOH is available if theres no freec
  colnames(ascat_loh)[3:4]=c('start','end')
  cnvs=ascat_cnv[nMajor+nMinor != 2] # this removes cnLOH
  cnvs$length=paste0(round((cnvs$endpos-cnvs$startpos)/1e6,2),'M')
  cnvs$rank_score <- 0
  cnvs$rank_terms <- ''
  cnvs$cancer_genes <- ''
  
  for (i in which(cnvs$endpos-cnvs$startpos < 50e6)) {
    genes=tumorgenes[chr==cnvs$chr[i] & start<cnvs$end[i] & end>cnvs$start[i],`Gene Symbol`] # genes in segment
    if (length(genes)>0) { # if any of our cancer genes
      cnvs$cancer_genes[i] <- paste(genes,collapse = ' ') # put in table
      if (any(genes %in% alltier1,na.rm = T)) {
        cnvs$rank_score[i]=2
        cnvs$rank_terms[i]='T1_gene'
      } else if (any(genes %in% alltier2,na.rm = T)) {
        cnvs$rank_score[i]=2
        cnvs$rank_terms[i]='T2_gene'
      }
    }
    
    # if the effect is focal
    if (cnvs[i,endpos-startpos]<3e6) {
      cnvs$rank_score[i]=cnvs$rank_score[i]+1
      cnvs$rank_terms[i]=paste(cnvs$rank_terms[i],'focal')
    }
    
    
    
    # if the effect is loss/gain/amp etc
    if (cnvs[i,nMajor+nMinor]==0) {
      cnvs$rank_score[i]=cnvs$rank_score[i]+2
      cnvs$rank_terms[i]=paste(cnvs$rank_terms[i],'hz_loss')
    } else if (cnvs[i,nMajor+nMinor] > 5) {
      cnvs$rank_score[i]=cnvs$rank_score[i]+2
      cnvs$rank_terms[i]=paste(cnvs$rank_terms[i],'high_amp')
    } else if (cnvs[i,nMajor+nMinor] > 2) {
      cnvs$rank_score[i]=cnvs$rank_score[i]+1
      cnvs$rank_terms[i]=paste(cnvs$rank_terms[i],'dup')
    } else if (cnvs[i,nMajor+nMinor] < 2) {
      cnvs$rank_score[i]=cnvs$rank_score[i]+1
      cnvs$rank_terms[i]=paste(cnvs$rank_terms[i],'del')
    }
  }
  
  ascat_cnv=cnvs[order(rank_score,decreasing = T)]
  if (write_tables) fwrite(ascat_cnv,file=paste0(sampleData$name,'_ascat_cnv.csv'))

  tableWrapper(ascat_cnv)
}
```



### View ControlFreec

#### Genome view
<!-- Plot Control Freec tumor copy number log ratio, tBAF and nBAF -->
```{r plot_control_freec,fig.width=15,fig.height=7}
if (project=='BTB') if (exists('tratio')) for (sample in unique(tratio$sample)) {
  cat('Control Freec: ',sample)
  par(mai=c(0,4,0,2))
  temp <- tratio[sample]
  temp$Ratio[temp$Ratio>3]=3
  
  g=ggplot()+ylab('B allele ratio and Coverage Ratio')+xlab(sample)+
    scale_color_gradientn(colours = c('violet','blue','black','orange','red'))+
    expand_limits(x=c(0,3200e6))+
    scale_y_continuous(limits=c(-1,3),
                       breaks = 0:2,
                       minor_breaks = seq(0,2.5,0.5),
                       labels = 0:2
    )
  # normal logR below
  g=g+geom_point(aes(x=cumstart,y=Ratio-0.3),col='darkgrey',data=nratio,alpha=1/5,shape='.')
  # tumor logR
  g=g+geom_point(aes(x=cumstart,y=Ratio,col=CopyNumber),data=temp,alpha=1/5,shape='.')
  # add smoothed
  #g=g+geom_line(aes(x=cumpos,y=tratio),stat="identity",colour="green",size=0.05,data=binned[sample])
  
  g=g+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
            axis.title.x=element_blank(),
            axis.text.x=element_blank(),
            axis.ticks.x=element_blank())
  # add tBAF  <--- downsampled by 10
  g=g+geom_point(data = tbaf[seq(1,nrow(tbaf),10)][sample],aes(x=cumstart,y=-0.5+0.5*BAF),
                 alpha=1/10,shape='.')
  # add nBAF  <--- downsampled by 10
  g=g+geom_point(data = nbaf[seq(1,nrow(nbaf),10)],aes(x=cumstart,y=-1+0.5*BAF),
                 alpha=1/10,shape='.')
  # add chromosome lines
  g=g+geom_segment(mapping = aes(x = starts,xend=starts,y= -1,yend=3),data=chrsz,inherit.aes = F,alpha=1/5)
  # add chromosome labels
  g=g+geom_text(aes(x=starts+0.5*length,y=0.1,label=label),data = chrsz,inherit.aes = F)
  
  print(g)
}
```

&nbsp;

#### Scatter plot
<!-- Scatter plot -->
```{r scatter_plot_control_freec, echo=FALSE,fig.width=15,fig.height=8,warning=F}
if (exists('binned')) for (sample in unique(binned$sample)) {
  cat('Control Freec: ',sample)
  binned$tmaf[binned$tmaf<0]=1
  g=ggplot()+expand_limits(y=c(.5,1))+expand_limits(x=c(.3,3))+
    geom_point(aes(x = tratio,y = tmaf,col=chr),data=binned[sample][tratio<3])+
    xlab('Tumor (1Mb mean) DNA ratio')+ylab('Tumor (1Mb mean) major allele ratio')+
    scale_y_continuous(breaks = seq(0.5,1,0.1)) +
    scale_x_continuous(breaks = seq(0.5,2.5,0.1))+
    theme(panel.grid.minor.x = element_blank(),
          panel.grid.major.x = element_line(colour="grey", size=0.2),
          panel.grid.major.y = element_line(colour="grey", size=0.2),
          panel.grid.minor.y = element_blank()
    )
  print(g)
}
```



### View Ascat {.active}

#### Genome view
<!-- Plot ASCAT tumor copy number log ratio, tBAF and nBAF -->
```{r plot_ascat,fig.width=15,fig.height=7}
if (exists('ascat_tratio')) for (sample in unique(ascat_tratio$sample)) {
  cat('ASCAT: ',sample)
  par(mai=c(0,4,0,2))
  temp <- ascat_tratio[sample]
  temp$smoothed[temp$smoothed>3]=3
  g=ggplot()+ylab('B allele ratio and Coverage Ratio')+
    scale_color_gradientn(colours = c('violet','blue','black','orange','red'))+
    scale_y_continuous(limits=c(-1,3),
                       breaks = 0:2,
                       minor_breaks = seq(0,2.5,0.5),
                       labels = 0:2
    )+
    expand_limits(x=c(0,3200e6))
  # using smoothed logR   <--- downsampled by 10
  g=g+geom_point(aes(x=cumstart,y=smoothed,color=CopyNumber),data=temp[c(1,seq(2,nrow(temp),10))],alpha=1/5,shape='.')
  # add MB-smoothed
  #g=g+geom_line(aes(x=cumpos,y=tratio),stat="identity",colour="green",size=0.05,data=ascat_binned[sample])
  
  g=g+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
            axis.title.x=element_blank(),
            axis.text.x=element_blank(),
            axis.ticks.x=element_blank())
  # add tBAF  <--- downsampled by 10
  g=g+geom_point(mapping = aes(x=cumstart,y=-0.5+0.5*BAF),data = ascat_tbaf[seq(1,nrow(ascat_tbaf),10)][sample],
                 alpha=1/10,shape='.')
  # add nBAF  <--- downsampled by 10
  g=g+geom_point(mapping = aes(x=cumstart,y=-1+0.5*BAF),data = ascat_nbaf[seq(1,nrow(ascat_nbaf),10)],
                 alpha=1/10,shape='.')
  # add chromosome lines
  g=g+geom_segment(mapping = aes(x = starts,xend=starts,y= -1,yend=3),data=chrsz,inherit.aes = F,alpha=1/5)
  # add chromosome labels
  g=g+geom_text(aes(x=starts+0.5*length,y=0.1,label=label),data = chrsz,inherit.aes = F)
  
  print(g)
}
```

&nbsp;

#### Scatter plot
<!-- Scatter plot -->
```{r scatter_plot_ascat, echo=FALSE,fig.width=15,fig.height=8,warning=F}
if (exists('ascat_binned')) for (sample in unique(ascat_binned$sample)) {
  cat('ASCAT: ',sample)
  g=ggplot()+expand_limits(y=c(.5,1))+expand_limits(x=c(.3,3))+
    geom_point(aes(x = tratio,y = tmaf,col=chr),data=ascat_binned[sample][tratio<3])+
    xlab('Tumor (1Mb mean) DNA ratio')+ylab('Tumor (1Mb mean) major allele ratio')+
    scale_y_continuous(breaks = seq(0.5,1,0.1)) +
    scale_x_continuous(breaks = seq(0.5,2.5,0.1))+
    theme(panel.grid.minor.x = element_blank(),
          panel.grid.major.x = element_line(colour="grey", size=0.2),
          panel.grid.major.y = element_line(colour="grey", size=0.2),
          panel.grid.minor.y = element_blank()
    )
  print(g)
}
```


## Structural variants {.tabset}

<!-- Read the tumor structural variant data -->
### Manta (tumor)
```{r load_manta_tumor, echo=FALSE}
manta_tumor_table=NULL
# first collect PASS ids from all samples
allpass=NULL
if (length(manta_tumor_file)>0) for (s in 1:length(manta_tumor_file)) {
  vcf=readVcf(file = manta_tumor_file[s],genome = reference_genome)
  pass=rowRanges(vcf)$FILTER=='PASS'
  allpass=c(allpass,names(vcf)[pass])
}
# then collect variants...
if (length(manta_tumor_file)>0) for (s in 1:length(manta_tumor_file)) {
  sample=strsplit(basename(manta_tumor_file[s]),'[.]')[[1]][1]
  cat(manta_tumor_file[s],'\n')
  
  vcf=readVcf(file = manta_tumor_file[s],genome = reference_genome)
  vcf=vcf[names(vcf) %in% allpass]
  
  
  if (length(vcf)>0) {
    
    g=geno(vcf)
    inf=info(vcf)
    rr=rowRanges(vcf)
    
    # Read counts
    srtable=as.data.table(g$SR)
    colnames(srtable)=paste0('SplitReadSupport_',c('N','T')) # Verified on a test sample
    prtable=as.data.table(g$PR)
    colnames(prtable)=paste0('PairedReadSupport_',c('N','T'))
    
    # Calculate allele ratio
    temp=data.table(
      pr.ref=sapply(prtable$PairedReadSupport_T,"[",1),
      sr.ref=sapply(srtable$SplitReadSupport_T,"[",1),
      pr.alt=sapply(prtable$PairedReadSupport_T,"[",2),
      sr.alt=sapply(srtable$SplitReadSupport_T,"[",2)
    )
    AFreq=apply(temp[,3:4],1,sum,na.rm=T)/
      (apply(temp[,3:4],1,sum,na.rm=T)+apply(temp[,1:2],1,sum,na.rm=T))
    
    
    # make table of variants
    sv_table=cbind(
      data.table(ID=as.character(names(vcf)),
                 sample,
                 chr=as.character(seqnames(rr))),
      as.data.table(rr)[,-1],
      AFreq,prtable,srtable,
      data.table(Swegen_count=rep(NA,length(vcf)),
                 rank_score=0,
                 rank_terms='',
                 LOH=''),
      as.data.table(inf)
    )
    sv_table$cumstart=sv_table$start
    sv_table$cumend=sv_table$end
    sv_table$altchr=sv_table$chr
    sv_table$altpos=sv_table$END
    sv_table$plot=F
    sv_table$arc= -1
    
    
    # Filter out variants seen 2+ (?) times in reference data
    ## Key has only chr,start,end
    key=sv_table[,c('chr','start','end')]
    if (project=='BTB') key$chr=substr(key$chr,4,6)
    key$imprecise='(pr)'
    ## If imprecise, round the pos to 10
    ix=sv_table$IMPRECISE==T
    key$imprecise[ix]='(impr)'
    key$start[ix]=round(key$start[ix]/10)*10
    key$end[ix]=round(key$end[ix]/10)*10
    key=paste(key$chr,key$start,key$end,key$imprecise)
    
    # put in Swegen count
    sv_table$Swegen_count=swegen_manta_all[key,value]
    sv_table$Swegen_count[is.na(sv_table$Swegen_count)]=0
    ## do the filter
    sv_table <- sv_table[Swegen_count<2]
    
  }
  
   if (exists('sv_table')) if (nrow(sv_table)>0) {
    # loop through all and extract endpoint chr and pos   <------ This one must be remade..
    for (i in 1:nrow(sv_table)) try( {                    # <----  sometimes error here, so try..
      t=strsplit(x = sv_table$ALT[[i]],split = ':')[[1]]
      if (length(t)>1 & t[1]!="<DUP") {
        tchr=str_extract(t[1],'[0-9,X,Y]*$')
        #if (is.na(tchr)) tchr=strsplit(x = t[1],split = 'CHR')[[1]][2]
        if (project=='BTB') sv_table$altchr[i] <- paste0('chr',tchr) else sv_table$altchr[i] <- tchr
        tt=str_extract(t[2],'^[0-9]*')
        sv_table$altpos[i]=as.numeric(tt)
      }
    }, silent=T)
    # for each chromosome get cumulative pos
    sv_table$altcumpos=sv_table$altpos
    for (i in 1:nrow(chrsz)) {
      ix=sv_table$chr==chrsz$chr[i]
      if (sum(ix)>0) {
        sv_table$cumstart[ix]=sv_table$start[ix]+chrsz$starts[i]
        sv_table$cumend[ix]=sv_table$end[ix]+chrsz$starts[i]
      }
      ix=sv_table$altchr==chrsz$chr[i]
      if (sum(ix)>0) {
        sv_table$altcumpos[ix]=sv_table$altpos[ix]+chrsz$starts[i]
      }
    }
    # decide how it is to be plotted (not represented elsewhere, up or down arc)
    for (i in 1:nrow(sv_table)) {
      if (sv_table$chr[i]==sv_table$altchr[i]) {
        # intrachromosomal: plot always, with "positive" arc
        sv_table$plot[i]=T
        sv_table$arc[i]=1 
      } else if (sv_table$altcumpos[i] > sv_table$cumstart[i]) {
        # interchromosomal: plot if mate is to right (else mate will be plotted) with negative arc
        sv_table$plot[i]=T 
        sv_table$arc[i]=-1 
      }
    }
    
    # snpEff annotation is in the ANN column.
    h=strsplit(info(header(vcf))['ANN',][,3],'annotations: \'')[[1]][2]
    snpEff_header=trimws(strsplit(h,'\\|')[[1]])
    
    # snpEff annotation is put in snpEff_table
    snpEff_table=matrix(data = NA,nrow = length(unlist(sv_table$ANN)),ncol = length(snpEff_header)+1)
    colnames(snpEff_table)=c('ID',snpEff_header)
    row=1
    for (i in 1:nrow(sv_table)) { # for each variant
      # for each VEP annotation:
      for (j in 1:length(sv_table$ANN[[i]]))  if (length(sv_table$ANN[[i]])>0) {
        line=strsplit(sv_table$ANN[[i]][j],'\\|')[[1]]
        snpEff_table[row,1]=sv_table$ID[i]
        snpEff_table[row,1+(1:length(line))]=line
        row=row+1
      }
    }
    snpEff_table=unique(as.data.table(snpEff_table))
    
    # Filter out annotations of certain types where the ID has >N annotations of that type
    ids=unique(snpEff_table$ID)
    for (id in ids) {
      common=c('protein_protein_contact', 'duplication', 'structural_interaction_variant', 'inversion',
          'transcript_ablation', 'feature_ablation', 'sequence_feature', 
          'intergenic_region', 'downstream_gene_variant', 'upstream_gene_variant')
      annotations=snpEff_table[ID==id,Annotation] # the annotations of this variant
      table=table(annotations[annotations %in% common])
      table=table[table>20] # the common annotations that appear >N times for this variant
      if (length(table)>0) {
        remove=which(snpEff_table$ID==id & snpEff_table$Annotation %in% names(table))
        snpEff_table=snpEff_table[-remove[-1],] # saves one to make sure the variant has some annotation
      }
    }
    
    # Add to data (all samples, one table)
    manta_tumor_table=rbind(manta_tumor_table,merge(sv_table,snpEff_table,by='ID',all=F))
    setkey(manta_tumor_table,'sample')
  }
}# done parsing each sample
  

# Prepare ranking and (some more) filtering
if (!is.null(manta_tumor_table)) if (nrow(manta_tumor_table)>0) {
  
  # ## Make table with most important info for ranking, and report
  # selected <- unique(manta_tumor_table[,.(ID,sample,SVTYPE,chr,start,REF,ALT,AFreq,PairedReadSupport_T,SplitReadSupport_T,
  #                        Swegen_count,Rank_score='',Rank_terms='',Gene_Name,Annotation,Annotation_Impact)])
  selection=manta_tumor_table[!is.na(Annotation)]
  
  if (nrow(selection)>0) {
    # Known cancer genes affect ranking by +1
    for (gene in unique(selection$Gene_Name)) if (!is.na(gene)) if (gene!='') {
      ix=selection$Gene_Name==gene
      gene=sort(strsplit(gene,'&')[[1]])
      # cosmic/local Tier2:
      if (any(gene %in% alltier2)) {
        selection$rank_score[ix]=2
        selection$rank_terms[ix]='T2_gene'
      }
      # tier 1 top priority:
      if (any(gene %in% alltier1)) {
        selection$rank_score[ix]=2
        selection$rank_terms[ix]='T1_gene'
      }
    }
    
    
    # Add high impact
    ix=selection$Annotation_Impact=='HIGH'
    if (any(ix)) {
      selection$rank_score[ix]=selection$rank_score[ix]+2
      selection$rank_terms[ix]=paste(selection$rank_terms[ix],'high_impact')
    }
    # Add moderate impact
    ix=selection$Annotation_Impact=='MODERATE'
    if (any(ix)) {
      selection$rank_score[ix]=selection$rank_score[ix]+1
      selection$rank_terms[ix]=paste(selection$rank_terms[ix],'moderate_impact')
    }
    
    
    # +1 for focal
    ix=selection$chr==selection$altchr & selection$end-selection$start < 3e6 & selection$altpos-selection$start < 3e6
    if (any(ix)) {
      selection$rank_score[ix]=selection$rank_score[ix]+1
      selection$rank_terms[ix]=paste(selection$rank_terms[ix],'focal')
    }
    
    # cosmic_>xx, fusion_gene or just fusion
    ix=grep('fusion',selection$Annotation)
    if (length(ix)>0) for (i in ix) if (selection$Gene_Name[i]!='') {
      gene=sort(strsplit(selection$Gene_Name[i],'&')[[1]])
      if (paste(gene,collapse = ' ') %in% cosmic_fusions[value>5,name]) {
        selection$rank_score[i]=selection$rank_score[i]+3
        selection$rank_terms[i]=paste(selection$rank_terms[i],'cosmic_>5')
      } else if (paste(gene,collapse = ' ') %in% cosmic_fusions$name) {
        selection$rank_score[i]=selection$rank_score[i]+2
        selection$rank_terms[i]=paste(selection$rank_terms[i],'cosmic_>1')
      }
      if (paste(gene,collapse = ' ') %in% allfusionpairs) {
        selection$rank_score[i]=selection$rank_score[i]+2
        selection$rank_terms[i]=paste(selection$rank_terms[i],'CGC_fusion')
      } else if (any(gene %in% allfusion)) {
        selection$rank_score[i]=selection$rank_score[i]+1
        selection$rank_terms[i]=paste(selection$rank_terms[i],'partial_CGC_fusion')
      } else {
        selection$rank_score[i]=selection$rank_score[i]+0
        selection$rank_terms[i]=paste(selection$rank_terms[i],'fusion')
      }
    }
    
    # -1 for ablation if long del
    ix=intersect(
      which(selection$SVTYPE=='DEL' & selection$chr==selection$altchr & abs(selection$altpos-selection$start) > 3e6),
      grep('ablation',selection$Annotation)
    )
    if (length(ix)>0) {
      selection$rank_score[ix]=selection$rank_score[ix]-1
      selection$rank_terms[ix]=paste(selection$rank_terms[ix],'long_del')
    }
    
    # -1 for duplication if long dup
    ix=intersect(
      which(selection$SVTYPE=='DUP' & selection$chr==selection$altchr & abs(selection$altpos-selection$start) > 10e6),
      grep('duplication',selection$Annotation)
    )
    if (length(ix)>0) {
      selection$rank_score[ix]=selection$rank_score[ix]-1
      selection$rank_terms[ix]=paste(selection$rank_terms[ix],'long_dup')
    }
 
    firstcols=c('ID','sample','Gene_Name','rank_score','rank_terms','LOH','AFreq','Annotation','Annotation_Impact','Swegen_count')
    cols=colnames(selection)
    setcolorder(x = selection,neworder = c(firstcols,cols[!cols %in% firstcols]))
    
    manta_tumor_selected <- selection[order(Feature_ID)][order(rank_score,decreasing = T)]
    if (write_tables) fwrite(manta_tumor_selected[,-c('ANN','CSQ','Gene_ID')][rank_score>1],file=paste0(sampleData$name,'_manta_tumor.csv'))
    tableWrapper(manta_tumor_selected[,-c('ANN','CSQ','Gene_ID')][rank_score>1])
  } 
}


```

<!-- Read the normal structural variant data -->
### Manta (normal)
```{r load_manta_normal, echo=FALSE}
if (project=='BTB') {
  manta_normal_table=NULL
  # first collect PASS ids from all samples
  allpass=NULL
  vcfs=list()
  if (length(manta_normal_file)>0) for (s in 1:length(manta_normal_file)) {
    vcfs[[manta_normal_file[s]]]=readVcf(file = manta_normal_file[s],genome = reference_genome)
    pass=rowRanges(vcfs[[manta_normal_file[s]]])$FILTER=='PASS'
    allpass=c(allpass,names(vcfs[[manta_normal_file[s]]])[pass])
  }
  # then collect variants...
  if (length(manta_normal_file)>0) for (s in 1:length(manta_normal_file)) {
    sample=strsplit(basename(manta_normal_file[s]),'[.]')[[1]][1]
    cat(manta_normal_file[s],'\n')
    
    vcf=vcfs[[manta_normal_file[s]]]
    vcf=vcf[names(vcf) %in% allpass]

    if (length(vcf)>0) {
      
      
      g=geno(vcf)
      inf=info(vcf)
      rr=rowRanges(vcf)
      
      # Read counts
      srtable=as.data.table(g$SR)
      colnames(srtable)=paste0('SplitReadSupport_',c('N')) # Verified on a test sample
      prtable=as.data.table(g$PR)
      colnames(prtable)=paste0('PairedReadSupport_',c('N'))
      
      # Calculate allele ratio
      temp=data.table(
        pr.ref=sapply(prtable$PairedReadSupport_N,"[",1),
        sr.ref=sapply(srtable$SplitReadSupport_N,"[",1),
        pr.alt=sapply(prtable$PairedReadSupport_N,"[",2),
        sr.alt=sapply(srtable$SplitReadSupport_N,"[",2)
      )
      AFreq=apply(temp[,3:4],1,sum,na.rm=T)/
        (apply(temp[,3:4],1,sum,na.rm=T)+apply(temp[,1:2],1,sum,na.rm=T))
      
      # make table of variants
      sv_table=cbind(
        data.table(ID=as.character(names(vcf)),
                   sample,
                   chr=as.character(seqnames(rr))),
        as.data.table(rr)[,-1],
        AFreq,prtable,srtable,
        data.table(Swegen_count=rep(NA,length(vcf)),
                   rank_score=0,
                   rank_terms='',
                   LOH=''),
        as.data.table(inf)
      )
      sv_table$cumstart=sv_table$start
      sv_table$cumend=sv_table$end
      sv_table$altchr=sv_table$chr
      sv_table$altpos=sv_table$END
      sv_table$plot=F
      sv_table$arc= -1
      
      # Filter out variants seen 2+ (?) times in reference data
      ## Key has only chr,start,end
      key=sv_table[,c('chr','start','end')]
      if (project=='BTB') key$chr=substr(key$chr,4,6)
      key$imprecise='(pr)'
      ## If imprecise, round the pos to 10
      ix=sv_table$IMPRECISE==T
      key$imprecise[ix]='(impr)'
      key$start[ix]=round(key$start[ix]/10)*10
      key$end[ix]=round(key$end[ix]/10)*10
      key=paste(key$chr,key$start,key$end,key$imprecise)
      
      # put in Swegen count
      sv_table$Swegen_count=swegen_manta_all[key,value]
      sv_table$Swegen_count[is.na(sv_table$Swegen_count)]=0
      ## do the filter
      sv_table <- sv_table[Swegen_count<2]
    }
      
    if (exists('sv_table')) if (nrow(sv_table)>0) {
      # loop through all and extract endpoint chr and pos   <------ This one must be remade..
      for (i in 1:nrow(sv_table)) try( {                    # <----  sometimes error here, so try..
        t=strsplit(x = sv_table$ALT[[i]],split = ':')[[1]]
        if (length(t)>1 & t[1]!="<DUP") {
          tchr=str_extract(t[1],'[0-9,X,Y]*$')
          #if (is.na(tchr)) tchr=strsplit(x = t[1],split = 'CHR')[[1]][2]
          if (project=='BTB') sv_table$altchr[i] <- paste0('chr',tchr) else sv_table$altchr[i] <- tchr
          tt=str_extract(t[2],'^[0-9]*')
          sv_table$altpos[i]=as.numeric(tt)
        }
      }, silent=T)
      # for each chromosome get cumulative pos
      sv_table$altcumpos=sv_table$altpos
      for (i in 1:nrow(chrsz)) {
        ix=sv_table$chr==chrsz$chr[i]
        if (sum(ix)>0) {
          sv_table$cumstart[ix]=sv_table$start[ix]+chrsz$starts[i]
          sv_table$cumend[ix]=sv_table$end[ix]+chrsz$starts[i]
        }
        ix=sv_table$altchr==chrsz$chr[i]
        if (sum(ix)>0) {
          sv_table$altcumpos[ix]=sv_table$altpos[ix]+chrsz$starts[i]
        }
      }
      # decide how it is to be plotted (not represented elsewhere, up or down arc)
      for (i in 1:nrow(sv_table)) {
        if (sv_table$chr[i]==sv_table$altchr[i]) {
          # intrachromosomal: plot always, with "positive" arc
          sv_table$plot[i]=T
          sv_table$arc[i]=1 
        } else if (sv_table$altcumpos[i] > sv_table$cumstart[i]) {
          # interchromosomal: plot if mate is to right (else mate will be plotted) with negative arc
          sv_table$plot[i]=T 
          sv_table$arc[i]=-1 
        }
      }
      
      # snpEff annotation is in the ANN column.
      h=strsplit(info(header(vcf))['ANN',][,3],'annotations: \'')[[1]][2]
      snpEff_header=trimws(strsplit(h,'\\|')[[1]])
      
      # snpEff annotation is put in snpEff_table
      snpEff_table=matrix(data = NA,nrow = length(unlist(sv_table$ANN)),ncol = length(snpEff_header)+1)
      colnames(snpEff_table)=c('ID',snpEff_header)
      row=1
      for (i in 1:nrow(sv_table)) { # for each variant
        # for each VEP annotation:
        for (j in 1:length(sv_table$ANN[[i]]))  if (length(sv_table$ANN[[i]])>0) {
          line=strsplit(sv_table$ANN[[i]][j],'\\|')[[1]]
          snpEff_table[row,1]=sv_table$ID[i]
          snpEff_table[row,1+(1:length(line))]=line
          row=row+1
        }
      }
      snpEff_table=unique(as.data.table(snpEff_table))[Annotation_Impact=='HIGH']  # <--- only HIGH impact kept for the normal
      
      # Filter out annotations of certain types where the ID has >N annotations of that type
      ids=unique(snpEff_table$ID)
      for (id in ids) {
        common=c('protein_protein_contact', 'duplication', 'structural_interaction_variant', 'inversion',
                 'transcript_ablation', 'feature_ablation', 'sequence_feature', 
                 'intergenic_region', 'downstream_gene_variant', 'upstream_gene_variant')
        annotations=snpEff_table$Annotation[snpEff_table$ID==id] # the annotations of this variant
        table=table(annotations[annotations %in% common])
        table=table[table>20] # the common annotations that appear >N times for this variant
        if (length(table)>0) {
          #browser()
          remove=which(snpEff_table$ID==id & snpEff_table$Annotation %in% names(table))
          snpEff_table=snpEff_table[-remove[-1],] # saves one to make sure the variant has some annotation
        }
      }
      
      # Add to data (all samples, one table)
      manta_normal_table=rbind(manta_normal_table,merge(sv_table,snpEff_table,by='ID',all=F))
      setkey(manta_normal_table,'sample')
    }
  }# done parsing each sample
  
  
  # Prepare ranking and (some more) filtering
  if (!is.null(manta_normal_table)) if (nrow(manta_normal_table)>0) {
    
    # ## Make table with most important info for ranking, and report
    # selected <- unique(manta_normal_table[,.(ID,sample,SVTYPE,chr,start,REF,ALT,AFreq,PairedReadSupport_T,SplitReadSupport_T,
    #                        Swegen_count,Rank_score='',Rank_terms='',Gene_Name,Annotation,Annotation_Impact)])
    selection=manta_normal_table[!is.na(Annotation)]
    
    if (nrow(selection)>0) {
      # Known cancer genes affect ranking by +1
      for (gene in unique(selection$Gene_Name)) if (!is.na(gene)) if (gene!='') {
        ix=selection$Gene_Name==gene
        gene=sort(strsplit(gene,'&')[[1]])
        # cosmic/local Tier2:
        if (any(gene %in% alltier2)) {
          selection$rank_score[ix]=2
          selection$rank_terms[ix]='T2_gene'
        }
        # tier 1 top priority:
        if (any(gene %in% alltier1)) {
          selection$rank_score[ix]=2
          selection$rank_terms[ix]='T1_gene'
        }
      }
      
      
      # Add high impact
      ix=selection$Annotation_Impact=='HIGH'
      if (any(ix)) {
        selection$rank_score[ix]=selection$rank_score[ix]+2
        selection$rank_terms[ix]=paste(selection$rank_terms[ix],'high_impact')
      }
      # Add moderate impact
      ix=selection$Annotation_Impact=='MODERATE'
      if (any(ix)) {
        selection$rank_score[ix]=selection$rank_score[ix]+1
        selection$rank_terms[ix]=paste(selection$rank_terms[ix],'moderate_impact')
      }
      
      # +1 for focal
      ix=selection$chr==selection$altchr & selection$end-selection$start < 3e6 & selection$altpos-selection$start < 3e6
      if (any(ix)) {
        selection$rank_score[ix]=selection$rank_score[ix]+1
        selection$rank_terms[ix]=paste(selection$rank_terms[ix],'focal')
      }
      
      # cosmic_>xx, fusion_gene or just fusion
      ix=grep('fusion',selection$Annotation)
      if (length(ix)>0) for (i in ix) if (selection$Gene_Name[i]!='') {
        gene=sort(strsplit(selection$Gene_Name[i],'&')[[1]])
        if (paste(gene,collapse = ' ') %in% cosmic_fusions[value>5,name]) {
          selection$rank_score[i]=selection$rank_score[i]+3
          selection$rank_terms[i]=paste(selection$rank_terms[i],'cosmic_>5')
        } else if (paste(gene,collapse = ' ') %in% cosmic_fusions$name) {
          selection$rank_score[i]=selection$rank_score[i]+2
          selection$rank_terms[i]=paste(selection$rank_terms[i],'cosmic_>1')
        }
        if (paste(gene,collapse = ' ') %in% allfusionpairs) {
          selection$rank_score[i]=selection$rank_score[i]+2
          selection$rank_terms[i]=paste(selection$rank_terms[i],'CGC_fusion')
        } else if (any(gene %in% allfusion)) {
          selection$rank_score[i]=selection$rank_score[i]+1
          selection$rank_terms[i]=paste(selection$rank_terms[i],'partial_CGC_fusion')
        } else {
          selection$rank_score[i]=selection$rank_score[i]+0
          selection$rank_terms[i]=paste(selection$rank_terms[i],'fusion')
        }
      }
      
      # -1 for ablation if long del
      ix=intersect(
        which(selection$SVTYPE=='DEL' & selection$chr==selection$altchr & abs(selection$altpos-selection$start) > 3e6),
        grep('ablation',selection$Annotation)
      )
      if (length(ix)>0) {
        selection$rank_score[ix]=selection$rank_score[ix]-1
        selection$rank_terms[ix]=paste(selection$rank_terms[ix],'long_del')
      }
      
      # -1 for duplication if long dup
      ix=intersect(
        which(selection$SVTYPE=='DUP' & selection$chr==selection$altchr & abs(selection$altpos-selection$start) > 10e6),
        grep('duplication',selection$Annotation)
      )
      if (length(ix)>0) {
        selection$rank_score[ix]=selection$rank_score[ix]-1
        selection$rank_terms[ix]=paste(selection$rank_terms[ix],'long_dup')
      }
      
      # Add Control Freec LOH.
      try( {
        if (nrow(selection)>0 & !is.null(freec_loh)) for (i in 1:nrow(selection)) {
          ix=freec_loh$chr==selection$chr[i] &
            freec_loh$start<selection$start[i] &
            freec_loh$end>selection$end[i]
          ix=ix[!is.na(ix)]
          if (sum(ix) >0) selection$LOH[i]='Y'
        }
      },silent=T)
      
      firstcols=c('ID','sample','Gene_Name','rank_score','rank_terms','LOH','AFreq','Annotation','Annotation_Impact','Swegen_count')
      cols=colnames(selection)
      setcolorder(x = selection,neworder = c(firstcols,cols[!cols %in% firstcols]))
      
      
      manta_normal_selected <- selection[order(Feature_ID)][order(rank_score,decreasing = T)]
      if (write_tables) fwrite(manta_normal_selected[,-c('ANN','CSQ','Gene_ID')][rank_score>1],file=paste0(sampleData$name,'_manta_normal.csv'))
      tableWrapper(manta_normal_selected[,-c('ANN','CSQ','Gene_ID')][rank_score>1])
    } 
  }
}
```

&nbsp;


<!-- Plot the tumor and normal (also showing copy number) structural variant data -->
### View structural variants {.active}
```{r plot_manta, echo=FALSE,fig.width=15,fig.height=4}
try( if (exists('manta_tumor_selected')) {

  cat(manta_tumor_file,manta_normal_file,sep='\n')

  sv_table=unique(manta_tumor_selected[,.(sample,chr,altchr,cumstart,altcumpos,IMPRECISE,SVTYPE,plot)])
  setkey(sv_table,'sample')
  sv_samples=sort(unique(sv_table$sample))
  if (exists('tratio')) {
    cnv_table=tratio[seq(1,nrow(tratio),10),.(sample,x=cumstart,y=log2(Ratio))]
    cnv_samples=sort(unique(cnv_table$sample))
  } else if (exists('ascat_tratio')) {
    cnv_table=ascat_tratio[seq(1,nrow(ascat_tratio),100),.(sample,x=cumstart,y=log2(smoothed))]
    cnv_samples=sort(unique(cnv_table$sample))
  }
  
  for (s in 1:length(sv_samples)) {
    
    cat('Manta:',sv_samples[s])
    
    par(mai=c(0,4,0,2))
    g=ggplot()+ylab('Structural variants')+expand_limits(x=c(0,3210e6))+ylim(-1.5,1.5)+
      xlab(paste0('SVs: ',sv_samples[s],',  CNVs: ',cnv_samples[s]))+ylab(sv_samples[s])+
      scale_color_manual(values=c("BND"="#E41A1C", "DEL"="#377EB8", "INV"="#4DAF4A", "INS"="#984EA3", "DUP"="#FF7F00"))#+
      #scale_linetype_manual(values=c('TRUE'=2,'FALSE'=1))
    # fix theme
    g=g+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
              axis.title.x=element_blank(),
              axis.text.x=element_blank(),
              axis.ticks.x=element_blank(),
              axis.title.y=element_blank(),
              axis.text.y=element_blank(),
              axis.ticks.y=element_blank())
    # add chromosome lines
    g=g+geom_segment(mapping = aes(x = starts,xend=starts,y=-1.5,yend=1.5),data=chrsz,inherit.aes = F,alpha=1)
    # add copy data if present
    ### TUMOR LOGRATIO  
    if (exists('cnv_table')) {
      g=g+geom_line(aes(x=x,y=y),
                    data = cnv_table[cnv_samples[s]],alpha=1/3,
                    stat="identity",colour="black")
    }
    
    # add chromosome labels
    g=g+geom_text(aes(x=starts+0.5*length,y=-0.7,label=label),data = chrsz,inherit.aes = F)
    # add the TUMOR structural variants
    if (exists('manta_tumor_table')) {
      temp=manta_tumor_table[sv_samples[s]][plot==T][IMPRECISE==FALSE]
      if (nrow(temp)>0) g=g+geom_curve(aes(x = cumstart,xend = altcumpos,y=0.5,yend=0.5001,col=SVTYPE),#,linetype=IMPRECISE),
                                       data = temp,alpha=1/2,
                                       stat = "identity",curvature=-0.2,size=0.5,
                                       position = "identity", angle = 90, ncp = 10,
                                       arrow = arrow(length=unit(0.10,"cm"), ends="both", type = "open"), 
                                       lineend = "butt", na.rm = FALSE, show.legend = NA,
                                       inherit.aes = FALSE)
    }
  print(g)
  }
}, silent=T)
```



## Somatic point mutations {.tabset .active}

### Mutect 2
<!-- Read annotated small variant data (Mutect 2) -->
```{r load_mutect2, echo=FALSE}
# first collect PASS ids from all samples
allpass=NULL
if (exists('mutect2_file')) if (length(mutect2_file)>0) {
  for (s in 1:length(mutect2_file)) {
    vcf=readVcf(file = mutect2_file[s],genome = reference_genome)
    pass=rowRanges(vcf)$FILTER=='PASS'
    allpass=c(allpass,names(vcf)[pass])
  }
  # then collect variants...
  mutect2_table=NULL
  for (s in 1:length(mutect2_file)) {
    sample=strsplit(basename(mutect2_file[s]),'[.]')[[1]][1]

    vcf=readVcf(file = mutect2_file[s],genome = reference_genome)
    vcf=vcf[names(vcf) %in% allpass]
 
#     # pre-filtering by SWAF / Swegen_count to avoid extreme nummber of variants
#     if (project=='BTB') {
#       swaf=as.data.table(info(vcf)$SWAF)
#       remove_swaf=unique(swaf$group[which(swaf$value>=0.01)])
#       in_ref=snptable[names(vcf)]
#       remove_swegen=which(in_ref$value>=10)
#       vcf=vcf[-union(remove_swaf,remove_swegen)]
#       topmed=as.data.table(info(vcf)$TOPMED)
#       vcf=vcf[-which(topmed$value>=0.01)]
#     } else {
#       in_ref=snptable[names(vcf)]
#       remove_swegen=which(in_ref$value>=10)
#       vcf=vcf[-remove_swegen]
#     }

    #m2vcf <- vcf # for later export

    if (length(vcf)>0) {
      rr=rowRanges(vcf)
      g=geno(vcf)
      inf=info(vcf)

      # manipulate into a data frame with relevant data
      mutations=as.data.table(ranges(rr))
      mutations$chr <- as.character(seqnames(rr))
      mutations$sample=sample
      mutations=mutations[,.(ID=names,sample,chr,start,end,width)]

      mutations$ref=as.character(ref(vcf))
      mutations$alt=as.data.table(alt(vcf))[, .(values = list(value)), by = group][,values]
      mutations$type=paste0(mutations$ref,'>',mutations$alt)
      mutations$type[nchar(mutations$ref)>nchar(mutations$alt)]='del'
      mutations$type[nchar(mutations$ref)<nchar(mutations$alt)]='ins'
      mutations$type[nchar(mutations$type)>3]='other'
      mutations$type[mutations$type=='T>G']='A>C'
      mutations$type[mutations$type=='T>C']='A>G'
      mutations$type[mutations$type=='T>A']='A>T'
      mutations$type[mutations$type=='G>T']='C>A'
      mutations$type[mutations$type=='G>C']='C>G'
      mutations$type[mutations$type=='G>A']='C>T'

      # Headers are sometimes sample names instead of TUMOR-NORMAL
      headers=colnames(g$AD)
      if (!'TUMOR' %in% headers) {
        ix=grep('-02$',headers)
        if (length(ix)==0) ix=grep('[TR]',headers)
        headers[ix]='TUMOR'
        headers[-ix]='NORMAL'
        colnames(g$AD)=headers
        colnames(g$AF)=headers # <-- assumes the same header order in AF
      }
      mutations$AFreq=round(sapply(g$AF[,'TUMOR'], "[", 1),2)
      ad=as.data.table(g$AD)
      colnames(ad)=paste0('AD_',colnames(ad))
      mutations=cbind(mutations,ad)
      # mutations$DP=sapply(g$AD[,'TUMOR'], sum)
      # mutations$AD=sapply(g$AD[,'TUMOR'], "[[", 2)
      # mutations$DPn=sapply(g$AD[,'NORMAL'], sum)
      # mutations$ADn=sapply(g$AD[,'NORMAL'], "[[", 2)
      mutations$reads=''
      temp=sapply(g$AD[,'TUMOR'], "[[", 2)
      mutations$reads[temp<5]='<5'
      mutations$reads[temp>=5]='5'

      # add counts from the new sources
      suppressWarnings( {
        t=unlist(lapply(X=inf$CAF,FUN=max,na.rm=T))
        t[is.infinite(t)]=0
        mutations$CAF=t
        t=unlist(lapply(X=inf$SWAF,FUN=max,na.rm=T))
        t[is.infinite(t)]=0
        mutations$SWAF=t
        t=unlist(lapply(X=inf$TOPMED,FUN=max,na.rm=T))
        t[is.infinite(t)]=0
        mutations$TOPMED=t
      } )

      # Add Swegen counts
      mutations$Swegen_count=snptable[names(vcf),value]
      mutations$Swegen_count[is.na(mutations$Swegen_count)]=0
      # # Some have multiple rsIDs (diminishingly few)
      # ix=grep(';',mutations$names)
      # ids=data.table(ix,id=strsplit(mutations$names[ix],';'))
      # ids$counts=unlist(lapply(ids$id,function(x) return(max(snptable[x,value],na.rm=T))))
      # ids$counts[is.infinite(ids$counts)]=0
      # mutations$Swegen_count[ix]=ids$counts
      # Also double check Swegen in case pos in reference but rsID or not properly matched in data
      pos=paste0(mutations$chr,':',mutations$start,'_',mutations$ref,'/',mutations$alt)
      snps=snptable[c(pos)][!is.na(value)] # extract from reference[those present]
      if (nrow(snps)>0) mutations$Swegen_count[match(snps$name,pos)]=snps$value # put
      # # Some have multiple alt alleles (diminishingly few)
      # ix=grep(',',mutations$alt)
      # for (i in ix) {
      #   alts=mutations$alt[[i]]
      #   snps=snptable[paste0('chr',mutations$chr[i],':',mutations$start[i],'_',mutations$ref[i],'/',alts)]
      #   mutations$Swegen_count[i]=max(c(0,snps$value),na.rm = T)
      # }

      ## Annotate by cosmic (for possibly retaining non PASS hotspots, which is not necessary smart)
      key=paste0(substr(mutations$chr,4,6),':',mutations$start,'-',mutations$end)
      key=str_replace(key,'X:','23:')
      key=str_replace(key,'Y:','24:')
      
      counts=cbind(cosmic_coding[key,value],cosmic_noncoding[key,value],0)
      max_=apply(counts,1,max,na.rm=T)
      mutations$Cosmic_count=max_


      mutations$rank_score=0
      mutations$rank_terms=''
      mutations$LOH=''

      # Add Control Freec LOH.
      try( {
        if (nrow(mutations)>0 & !is.null(freec_loh)) for (i in 1:nrow(mutations)) {
          ix=freec_loh$chr==mutations$chr[i] &
            freec_loh$start<mutations$start[i] &
            freec_loh$end>mutations$end[i]
          ix=ix[!is.na(ix)]
          if (sum(ix) >0) mutations$LOH[i]='Y'
        }},silent=T)

      # "info"" annotations also need to be added
      mutations=cbind(mutations,as.data.table(info(vcf))[,.(CSQ)])

      # keep only PASS
      # mutations=mutations[filter=='PASS',]

      # get snpEff headers
      # snpeff_header=strsplit(info(header(vcf))['ANN',][,3],'ions: \'')[[1]][2]
      # snpeff_header=strsplit(snpeff_header,' \\| ')[[1]]
      ## get VEP headers
      vep_header=strsplit(info(header(vcf))['CSQ',][,3],'Format: ')[[1]][2]
      vep_header=strsplit(vep_header,'\\|')[[1]]


      # VEP annotation is put in annotation_table
      annotation_table=matrix(data = NA,nrow = length(unlist(mutations$CSQ)),ncol = length(vep_header)+1)
      colnames(annotation_table)=c('ID',vep_header)
      row=1
      for (i in 1:nrow(mutations)) { # for each variant
        # for each VEP annotation:
        for (j in 1:length(mutations$CSQ[[i]])) {
          line=strsplit(mutations$CSQ[[i]][j],'\\|')[[1]]
          annotation_table[row,1]=mutations$ID[i]
          annotation_table[row,1+(1:length(line))]=line
          row=row+1
        }
      }
      annotation_table=as.data.table(annotation_table)

      # Annotation table then concatenated
      mutect2_table=rbind(mutect2_table,merge(mutations,annotation_table,by='ID'))
    }
  } # done collecting from vcf files

  
  # # filtering by SWAF / Swegen_count (same as with Haplotypecaller)
  # if (project=='BTB') {
  #   mutect2_table=mutect2_table[-which(mutect2_table$SWAF>=0.01)]
  #   mutect2_table=mutect2_table[-which(mutect2_table$Swegen_count>=10)]
  #   mutect2_table=mutect2_table[-which(mutect2_table$TOPMED>=0.01)]
  # } else {
  #   mutect2_table=mutect2_table[-which(mutect2_table$Swegen_count>=10)]
  # }

  
  
  mutect2_table$cumstart=NA
  mutect2_table$cumend=NA
  # # for each chr get cumulative pos
  for (i in 1:nrow(chrsz)) {
    ix=mutect2_table$chr==chrsz$chr[i]
    mutect2_table$cumstart[ix]=mutect2_table$start[ix]+chrsz$starts[i]
    mutect2_table$cumend[ix]=mutect2_table$end[ix]+chrsz$starts[i]
  }
  setkey(mutect2_table,'sample')

  selection=mutect2_table[,-c('CSQ')] # not needed after parsing/merging the annotations

  if (nrow(selection)>0) {

    # Cosmic/local Tier2 :
    ix=selection$SYMBOL %in% alltier2
    selection$rank_score[ix]=2
    selection$rank_terms[ix]='T2_gene'
    # tier 1 priority:
    ix=selection$SYMBOL %in% alltier1
    selection$rank_score[ix]=2
    selection$rank_terms[ix]='T1_gene'

    # Add high impact
    ix=selection$IMPACT=='HIGH'
    if (any(ix)) {
      selection$rank_score[ix]=selection$rank_score[ix]+2
      selection$rank_terms[ix]=paste(selection$rank_terms[ix],'high_impact')
    }

    # Add moderate impact
    ix=selection$IMPACT=='MODERATE'
    if (any(ix)) {
      selection$rank_score[ix]=selection$rank_score[ix]+1
      selection$rank_terms[ix]=paste(selection$rank_terms[ix],'moderate_impact')
    }

    # additional +1 if high impact and TSG
    ix=selection$IMPACT=='HIGH' & selection$SYMBOL %in% alltsg
    if (any(ix)) {
      selection$rank_score[ix]=selection$rank_score[ix]+1
      selection$rank_terms[ix]=paste(selection$rank_terms[ix],'high+TSG')
    }

    # Add clinvar pathogenic
    ix=grep('pathogenic',selection$CLIN_SIG)
    if (length(ix)>0) {
      selection$rank_score[ix]=selection$rank_score[ix]+2
      selection$rank_terms[ix]=paste(selection$rank_terms[ix],'clinvar')
    }

    # Add Polyphen/SIFT damaging/deleterious
    ix=union(grep('damaging',selection$PolyPhen),grep('deleterious',selection$SIFT))
    if (length(ix)>0) {
      selection$rank_score[ix]=selection$rank_score[ix]+1
      selection$rank_terms[ix]=paste(selection$rank_terms[ix],'polyphen/SIFT')
    }

    # Add hotspots and near hotspots (within 2 residues of a hotspot)
    key=paste(selection$SYMBOL,
              str_replace(string = selection$Protein_position,
                          pattern = '/.*',replacement = ''))
    ix <-
      key %in% paste(hotspots_snv[,Hugo_Symbol],hotspots_snv[,Amino_Acid_Position]) |
      key %in% paste(hotspots_inframe[,Hugo_Symbol],hotspots_inframe[,Amino_Acid_Position])  ## Warning: Exact match used with inframes
    if (any(ix)) {
      selection$rank_score[ix]=selection$rank_score[ix]+2
      selection$rank_terms[ix]=paste(selection$rank_terms[ix],'hotspot')
    }
    ix = !ix & key %in% near_hotspots # the near_hotspot excludes those that were a hotspot
    if (any(ix)) {
      selection$rank_score[ix]=selection$rank_score[ix]+1
      selection$rank_terms[ix]=paste(selection$rank_terms[ix],'near_hotspot')
    }
    

    # Add cosmic counts
    ix=which(selection$Cosmic_count>50)
    if (length(ix)>0) {
      selection$rank_score[ix]=selection$rank_score[ix]+2
      selection$rank_terms[ix]=paste(selection$rank_terms[ix],'cosmic_>50')
    }
    ix=which(selection$Cosmic_count>5 & selection$Cosmic_count<=50)
    if (length(ix)>0) {
      selection$rank_score[ix]=selection$rank_score[ix]+1
      selection$rank_terms[ix]=paste(selection$rank_terms[ix],'cosmic_>5')
    }
    
    # Special case for TERT promoter
    ix=which(selection$SYMBOL=='TERT' & selection$Consequence=='5_prime_UTR_variant')
    if (length(ix)>0) {
      selection$rank_score[ix]=selection$rank_score[ix]+2
      selection$rank_terms[ix]=paste(selection$rank_terms[ix],'TERT_5\'UTR')
    }

    # Add TF binding variants near (100kb) cancer genes
    ix=grep('TF',selection$Consequence)
    if (length(ix)>0) {
      selection$rank_score[ix]=selection$rank_score[ix]+2
      selection$rank_terms[ix]=paste(selection$rank_terms[ix],'TFBS')
      for (i in 1:length(ix)) {
        near=which(selection$cumend[ix[i]] > (tumorgenes$cumstart-100e3) & selection$cumstart[ix[i]] < (tumorgenes$cumend + 100e3) &
                     !selection$SYMBOL[ix[i]] %in% alltumorgenes)
        if (length(near)>0) {
          genes=paste(tumorgenes$`Gene Symbol`[near],collapse = ',')
          selection$rank_score[ix[i]]=selection$rank_score[ix[i]]+2
          selection$rank_terms[ix[i]]=paste(selection$rank_terms[ix[i]],paste0('near_',genes))
        }
      }
    }

    ix=which(selection$CANONICAL!='YES')
    if (length(ix)>0) {
      selection$rank_score[ix]=selection$rank_score[ix]-0
      selection$rank_terms[ix]=paste(selection$rank_terms[ix],'not_canonical')
    }

  }

  firstcols=c('ID','sample','SYMBOL','rank_score','rank_terms','LOH','AFreq','Consequence','IMPACT','SWAF','TOPMED','Swegen_count')
  cols=colnames(selection)
  setcolorder(x = selection,neworder = c(firstcols,cols[!cols %in% firstcols]))
  selection <- selection[order(cumstart,Allele)][order(rank_score,decreasing = T)]
  
  # # Take some information to the VCF
  # selected_for_vcf=selection[, .(LOH=LOH,Swegen_count=Swegen_count,rank_terms=paste0(rank_score,':',SYMBOL,' ',trimws(rank_terms)),rank_score=rank_score), by = ID]
  # selected_for_vcf=selected_for_vcf[, .(LOH=LOH,Swegen_count=Swegen_count,rank_terms=paste(unique(rank_terms),collapse=','),rank_score=paste(unique(rank_score),collapse=',')), by=ID]
  # selected_for_vcf$rank_terms[selected_for_vcf$rank_score=='0']=''
  # 
  
  mutect2_selected <- selection
  
  if (write_tables) fwrite(mutect2_selected,file=paste0(sampleData$name,'_mutect2_tumor.csv'))


  tableWrapper(mutect2_selected[,-c('cumstart','cumend','DOMAINS')][rank_score>0])
}


```

### Strelka
<!-- Read annotated small variant data (Strelka, two files) -->
```{r load_strelka, echo=FALSE}
allpass=NULL
vcfs=list()
if (length(strelka_snv_file)>0  & length(strelka_indel_file)>0) for (s in 1:length(strelka_snv_file)) {
  # read snvs
  vcfs[[strelka_snv_file[s]]]=readVcf(file = strelka_snv_file[s],genome=reference_genome)
  pass=rowRanges(vcfs[[strelka_snv_file[s]]])$FILTER=='PASS'
  allpass=c(allpass,names(vcfs[[strelka_snv_file[s]]])[pass])
  # read indels
  vcfs[[strelka_indel_file[s]]]=readVcf(file = strelka_indel_file[s],genome=reference_genome)
  pass=rowRanges(vcfs[[strelka_indel_file[s]]])$FILTER=='PASS'
  allpass=c(allpass,names(vcfs[[strelka_indel_file[s]]])[pass])
}
strelka_table=NULL
if (length(strelka_snv_file)>0  & length(strelka_indel_file)>0) {
  for (s in 1:length(strelka_snv_file)) {
    sample=strsplit(basename(strelka_snv_file[s]),'_somatic')[[1]][1]
    # first snvs
    vcf=vcfs[[strelka_snv_file[s]]]
    vcf=vcf[names(vcf) %in% allpass] # only those with PASS in either sample
    if (!'TUMOR' %in% colnames(vcf)) colnames(vcf)=c('NORMAL','TUMOR') # assumption that normal comes first
    # Collect sample-unspecific data for the [PASS in any] IDs into a table
    g=geno(vcf)
    rr=rowRanges(vcf)
    inf=info(vcf)
    table_snvs=NULL
    csq=NULL
    if (length(vcf)>0) {
      table_snvs=data.table(ID=as.character(names(vcf)),
                            sample=sample,
                            chr=as.character(seqnames(rr)),
                            start=start(rr),
                            end=end(rr),
                            ref=as.data.table(rr$REF)$x,
                            alt=as.data.table(rr$ALT)$value,
                            AFreq=NA,AD=NA,DP=as.data.table(g$DP)$TUMOR,
                            AD_normal=NA,DP_normal=as.data.table(g$DP)$NORMAL)
      # Collect variant allele depths for SNVs:
      for (this_ref in unique(table_snvs$ref)) for (this_alt in unique(table_snvs[ref==this_ref,alt])) {
        which_ones=table_snvs$ref==this_ref & table_snvs$alt==this_alt
        refcounts=as.data.table(data.frame(g[[paste0(this_ref,'U')]]))
        altcounts=as.data.table(data.frame(g[[paste0(this_alt,'U')]]))
        table_snvs$AFreq[which_ones]=round(altcounts[which_ones,TUMOR.1] /
                                             (altcounts[which_ones,TUMOR.1]+refcounts[which_ones,TUMOR.1]),2)
        table_snvs$AD[which_ones]=altcounts[which_ones,TUMOR.1]
        table_snvs$AD_normal[which_ones]=altcounts[which_ones,NORMAL.1]
      }
      table_snvs$reads='5'
      table_snvs$reads[table_snvs$AD<5]='<5'
      # CAF,SWAF,TOPMED
      if (project=='BTB') suppressWarnings({
        t=unlist(lapply(X=inf$CAF,FUN=max,na.rm=T))
        t[is.infinite(t)]=0
        table_snvs$CAF=t
        t=unlist(lapply(X=inf$SWAF,FUN=max,na.rm=T))
        t[is.infinite(t)]=0
        table_snvs$SWAF=t
        t=unlist(lapply(X=inf$TOPMED,FUN=max,na.rm=T))
        t[is.infinite(t)]=0
        table_snvs$TOPMED=t
      } )
      
      # The snv annotations will be used later:
      csq=as.data.table(info(vcf))[,CSQ]
    }
    
    # Same procedure for indels
    vcf=vcfs[[strelka_indel_file[s]]]
    vcf=vcf[names(vcf) %in% allpass] # only those with PASS in either sample
    if (!'TUMOR' %in% colnames(vcf)) colnames(vcf)=c('NORMAL','TUMOR') # assumption that normal comes first
    # Collect sample-unspecific data for the [PASS in any] IDs into a table
    g=geno(vcf)
    rr=rowRanges(vcf)
    inf=info(vcf)
    table_indels=NULL
    if (length(vcf)>0) {
      table_indels=data.table(ID=as.character(names(vcf)),
                              sample=sample,
                              chr=as.character(seqnames(rr)),
                              start=start(rr),
                              end=end(rr),
                              ref=as.data.table(rr$REF)$x,
                              alt=as.data.table(rr$ALT)$value,
                              AFreq=NA,AD=NA,DP=as.data.table(g$DP)$TUMOR,
                              AD_normal=NA,DP_normal=as.data.table(g$DP)$NORMAL)
      refcounts=as.data.table(data.frame(g$TAR))
      altcounts=as.data.table(data.frame(g$TIR))
      table_indels$AFreq=round(altcounts$TUMOR.1 / (altcounts$TUMOR.1 + refcounts$TUMOR.1),2)
      table_indels$AD_normal=altcounts$NORMAL.1
      table_indels$AD=altcounts$TUMOR.1
      table_indels$reads='5'
      table_indels$reads[table_indels$AD<5]='<5'
      # CAF,SWAF,TOPMED
      if (project=='BTB') suppressWarnings({
        t=unlist(lapply(X=inf$CAF,FUN=max,na.rm=T))
        t[is.infinite(t)]=0
        table_indels$CAF=t
        t=unlist(lapply(X=inf$SWAF,FUN=max,na.rm=T))
        t[is.infinite(t)]=0
        table_indels$SWAF=t
        t=unlist(lapply(X=inf$TOPMED,FUN=max,na.rm=T))
        t[is.infinite(t)]=0
        table_indels$TOPMED=t
      })
      
      # indel annotations added to snv annotations
      csq=c(csq,as.data.table(info(vcf))[,CSQ])
    }

    # Concatenate the snvs and indels tables for this patient and sample
    table_both=rbind(table_snvs,table_indels)
    table_both$Swegen_count=NA
    table_both$Cosmic_count=NA
    table_both$rank_score=0
    table_both$rank_terms=''
    table_both$LOH=''

    # Variant type
    table_both$type=paste0(table_both$ref,'>',table_both$alt)
    table_both$type[nchar(table_both$ref)>nchar(table_both$alt)]='del'
    table_both$type[nchar(table_both$ref)<nchar(table_both$alt)]='ins'
    table_both$type[nchar(table_both$type)>3]='other'
    table_both$type[table_both$type=='T>G']='A>C'
    table_both$type[table_both$type=='T>C']='A>G'
    table_both$type[table_both$type=='T>A']='A>T'
    table_both$type[table_both$type=='G>T']='C>A'
    table_both$type[table_both$type=='G>C']='C>G'
    table_both$type[table_both$type=='G>A']='C>T'

    table_both$CSQ=csq

    # get VEP headers from a vcf object (assumed never to differ between snvs and indels)
    vep_header=strsplit(info(header(vcf))['CSQ',][,3],'Format: ')[[1]][2]
    vep_header=strsplit(vep_header,'\\|')[[1]]

    # VEP annotation is put in annotation_table
    annotation_table=matrix(data = NA,nrow = length(unlist(csq)),ncol = length(vep_header)+1)
    colnames(annotation_table)=c('ID',vep_header)
    row=1
    for (i in 1:length(csq)) { # for each variant
      # for each VEP annotation:
      for (j in 1:length(csq[[i]])) {
        line=strsplit(csq[[i]][j],'\\|')[[1]]
        annotation_table[row,1]=table_both$ID[i]
        annotation_table[row,1+(1:length(line))]=line
        row=row+1
      }
    }

    # Annotation table then merged with the "table_both"
    strelka_table=rbind(strelka_table,merge(table_both,as.data.table(annotation_table),by='ID'))
    setkey(strelka_table,'sample')
  } # done parsing each sample
  rm(vcfs)

  # Add Control Freec LOH.  <-- slow.
  try( {
    if (nrow(strelka_table)>0 & !is.null(freec_loh)) for (i in 1:nrow(strelka_table)) {
      ix=freec_loh$chr==strelka_table$chr[i] &
        freec_loh$start<strelka_table$start[i] &
        freec_loh$end>strelka_table$end[i]
      ix=ix[!is.na(ix)]
      if (sum(ix) >0) strelka_table$LOH[i]='Y'
    }},silent=T)

  # Add Swegen counts
  by_pos=snptable[strelka_table$ID,value] # if the variant is as 10:10001801_C/T in database
  by_name1=snptable[str_extract(strelka_table$Existing_variation,"^rs[0-9]+"),value] # if first rsid
  by_name2=snptable[str_extract(strelka_table$Existing_variation,"rs[0-9]+$"),value] # if last rsid (often both)
  d=data.table(by_pos,by_name1,by_name2,default=0)
  d$max=apply(X = d,MARGIN = 1,FUN = max,na.rm=T)
  strelka_table$Swegen_count=d$max

  
  # # filtering by SWAF / Swegen_count (same as with Haplotypecaller)
  # if (project=='BTB') {
  #   strelka_table=strelka_table[-which(strelka_table$SWAF>=0.01)]
  #   strelka_table=strelka_table[-which(strelka_table$Swegen_count>=10)]
  #   strelka_table=strelka_table[-which(strelka_table$TOPMED>=0.01)]
  # } else {
  #   strelka_table=strelka_table[-which(strelka_table$Swegen_count>=10)]
  # }
  
  
  # # Add Cosmic counts
  # my_ranges=GRanges(seqnames = strelka_table$chr,
  #                   ranges = IRanges(as.numeric(strelka_table$start),as.numeric(strelka_table$end)))
  # ix=match(my_ranges,ranges19)
  # strelka_table$Cosmic_count=ranges19$mcols.value[ix]
  # strelka_table$Cosmic_count[is.na(strelka_table$Cosmic_count)]=''
  key=paste0(substr(strelka_table$chr,4,6),':',strelka_table$start,'-',strelka_table$end)
  if (project!='BTB') key=paste0(strelka_table$chr,':',strelka_table$start,'-',strelka_table$end)
  key=str_replace(key,'X:','23:')
  key=str_replace(key,'Y:','24:')
  counts=cbind(cosmic_coding[key,value],cosmic_noncoding[key,value],0)
  max_=apply(counts,1,max,na.rm=T)
  strelka_table$Cosmic_count=max_

  strelka_table$cumstart=strelka_table$start
  strelka_table$cumend=strelka_table$end
  # # for each chr get cumulative pos
  for (i in 1:nrow(chrsz)) {
    ix=strelka_table$chr==chrsz$chr[i]
    strelka_table$cumstart[ix]=strelka_table$start[ix]+chrsz$starts[i]
    strelka_table$cumend[ix]=strelka_table$end[ix]+chrsz$starts[i]
  }

  selection=strelka_table[,-c('CSQ')] # not needed after parsing/merging the annotations

  if (nrow(selection)>0) {

    # Cosmic/local Tier2 :
    ix=selection$SYMBOL %in% alltier2
    selection$rank_score[ix]=2
    selection$rank_terms[ix]='T2_gene'
    # tier 1 priority:
    ix=selection$SYMBOL %in% alltier1
    selection$rank_score[ix]=2
    selection$rank_terms[ix]='T1_gene'

    # Add high impact
    ix=selection$IMPACT=='HIGH'
    if (any(ix)) {
      selection$rank_score[ix]=selection$rank_score[ix]+2
      selection$rank_terms[ix]=paste(selection$rank_terms[ix],'high_impact')
    }

    # Add moderate impact
    ix=selection$IMPACT=='MODERATE'
    if (any(ix)) {
      selection$rank_score[ix]=selection$rank_score[ix]+1
      selection$rank_terms[ix]=paste(selection$rank_terms[ix],'moderate_impact')
    }

    # additional +1 if high impact and TSG
    ix=selection$IMPACT=='HIGH' & selection$SYMBOL %in% alltsg
    if (any(ix)) {
      selection$rank_score[ix]=selection$rank_score[ix]+1
      selection$rank_terms[ix]=paste(selection$rank_terms[ix],'high+TSG')
    }

    # Add clinvar pathogenic
    ix=grep('pathogenic',selection$CLIN_SIG)
    if (length(ix)>0) {
      selection$rank_score[ix]=selection$rank_score[ix]+2
      selection$rank_terms[ix]=paste(selection$rank_terms[ix],'clinvar')
    }

    # Add Polyphen/SIFT damaging/deleterious
    ix=union(grep('damaging',selection$PolyPhen),grep('deleterious',selection$SIFT))
    if (length(ix)>0) {
      selection$rank_score[ix]=selection$rank_score[ix]+1
      selection$rank_terms[ix]=paste(selection$rank_terms[ix],'polyphen/SIFT')
    }

    # Add hotspots and near hotspots (within 2 residues of a hotspot)
    key=paste(selection$SYMBOL,
              str_replace(string = selection$Protein_position,
                          pattern = '/.*',replacement = ''))    
    ix <-
      key %in% paste(hotspots_snv[,Hugo_Symbol],hotspots_snv[,Amino_Acid_Position]) |
      key %in% paste(hotspots_inframe[,Hugo_Symbol],hotspots_inframe[,Amino_Acid_Position])  ## Warning: Exact match used with inframes
    if (any(ix)) {
      selection$rank_score[ix]=selection$rank_score[ix]+2
      selection$rank_terms[ix]=paste(selection$rank_terms[ix],'hotspot')
    }
    ix = !ix & key %in% near_hotspots # the near_hotspot excludes those that were a hotspot
    if (any(ix)) {
      selection$rank_score[ix]=selection$rank_score[ix]+1
      selection$rank_terms[ix]=paste(selection$rank_terms[ix],'near_hotspot')
    }

    # Add cosmic counts
    ix=which(selection$Cosmic_count>50)
    if (length(ix)>0) {
      selection$rank_score[ix]=selection$rank_score[ix]+2
      selection$rank_terms[ix]=paste(selection$rank_terms[ix],'cosmic_>50')
    }
    ix=which(selection$Cosmic_count>5 & selection$Cosmic_count<=50)
    if (length(ix)>0) {
      selection$rank_score[ix]=selection$rank_score[ix]+1
      selection$rank_terms[ix]=paste(selection$rank_terms[ix],'cosmic_>5')
    }

    # Special case for TERT promoter
    ix=which(selection$SYMBOL=='TERT' & selection$Consequence=='5_prime_UTR_variant')
    if (length(ix)>0) {
      selection$rank_score[ix]=selection$rank_score[ix]+2
      selection$rank_terms[ix]=paste(selection$rank_terms[ix],'TERT_5\'UTR')
    }

    # Add TF binding variants near (100kb) cancer genes
    ix=grep('TF',selection$Consequence)
    if (length(ix)>0) {
      selection$rank_score[ix]=selection$rank_score[ix]+2
      selection$rank_terms[ix]=paste(selection$rank_terms[ix],'TFBS')
      for (i in 1:length(ix)) {
        near=which(selection$cumend[ix[i]] > (tumorgenes$cumstart-100e3) & selection$cumstart[ix[i]] < (tumorgenes$cumend + 100e3) &
                     !selection$SYMBOL[ix[i]] %in% alltumorgenes)
        if (length(near)>0) {
          genes=paste(tumorgenes$`Gene Symbol`[near],collapse = ',')
          selection$rank_score[ix[i]]=selection$rank_score[ix[i]]+2
          selection$rank_terms[ix[i]]=paste(selection$rank_terms[ix[i]],paste0('near_',genes))
        }
      }
    }
    
    ix=which(selection$CANONICAL!='YES')
    if (length(ix)>0) {
      selection$rank_score[ix]=selection$rank_score[ix]-0
      selection$rank_terms[ix]=paste(selection$rank_terms[ix],'not_canonical')
    }

  }

  firstcols=c('ID','sample','SYMBOL','rank_score','rank_terms','LOH','AFreq','Consequence','IMPACT','SWAF','TOPMED','Swegen_count')
  cols=colnames(selection)
  setcolorder(x = selection,neworder = c(firstcols,cols[!cols %in% firstcols]))
  
  strelka_selected <- selection[order(cumstart,Allele)][order(rank_score,decreasing = T)]
  if (write_tables) fwrite(strelka_selected,file=paste0(sampleData$name,'_strelka_tumor.csv'))
  tableWrapper(strelka_selected[,-c('cumstart','cumend','DOMAINS')][rank_score>0])
}
  

```


<!-- Plot Mutect 2 mutations along the genome -->
### View Mutect 2

```{r plot_mutect2, echo=FALSE,fig.width=15,fig.height=4}
if (exists('mutect2_file')) cat(mutect2_file,sep = '\n')
if (exists('mutect2_table')) for (sample in sort(unique(mutect2_table$sample))) {
  
  
  par(mai=c(1,4,0,2))

  # Basic plot element
  g_basic=
    ggplot()+
    ylab('Somatic mutation allele ratio')+
    expand_limits(y=c(0,1))+
    expand_limits(x=c(0,3210e6))+
    scale_fill_manual(values = c("A>C"="#1B9E77","A>G"="#D95F02", "A>T"="#7570B3", "C>A"="#E7298A", "C>G"="#66A61E", "C>T"="#E6AB02", "del"="#A6761D", "ins"="#666666"))+
    scale_shape_manual(values=c("A>C"=21,"A>G"=21, "A>T"=21, "C>A"=21, "C>G"=21, "C>T"=21, "del"=24, "ins"=25))+
    scale_color_manual(values=c('<5'='lightgrey','5'='black'))+
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          axis.title.x=element_blank(),
          axis.text.x=element_blank(),
          axis.ticks.x=element_blank())+
    geom_segment(mapping = aes(x = starts,xend=starts,y=0,yend=1),data=chrsz,inherit.aes = F,alpha=1/5)+
    geom_text(aes(x=starts+0.5*length,y=0,label=label),data = chrsz,inherit.aes = F)


  # Mutect2 plot
  cat('Mutect2: ',sample)
  temp=unique(mutect2_table[sample][,.(cumstart,reads,AFreq,type,SYMBOL,Protein_position,Amino_acids,IMPACT)])
  g_mutect2=g_basic+
    xlab(paste('Mutect2: ',sample))+
    geom_point(data = temp,mapping = aes(x=cumstart,y=AFreq,fill=type,shape=type,col=reads),alpha=1)+
    geom_label_repel(data = temp[SYMBOL %in% alltier1 & IMPACT %in% c('MODERATE','HIGH'),],
                     mapping = aes(x=cumstart,y=AFreq,label=paste(SYMBOL,Protein_position,Amino_acids)),
                     inherit.aes = F,size=3,nudge_y = 0.2,col='black')+
    geom_label_repel(data = temp[SYMBOL %in% alltier2 & IMPACT %in% c('MODERATE','HIGH'),],
                     mapping = aes(x=cumstart,y=AFreq,label=paste(SYMBOL,Protein_position,Amino_acids)),
                     inherit.aes = F,size=3,nudge_y = 0.2,col='grey')
  print(g_mutect2)

}
```


<!-- Plot Strelka mutations along the genome -->
### View Strelka {.active}

```{r plot_strelka, echo=FALSE,fig.width=15,fig.height=4}
if (exists('strelka_indel_file')) cat(strelka_indel_file,strelka_snv_file,sep = '\n')
if (exists('strelka_table')) for (sample in sort(unique(strelka_table$sample))) {
  par(mai=c(1,4,0,2))

  # Basic plot element
  g_basic=
    ggplot()+
    ylab('Somatic mutation allele ratio')+
    expand_limits(y=c(0,1))+
    expand_limits(x=c(0,3210e6))+
    scale_fill_manual(values = c("A>C"="#1B9E77","A>G"="#D95F02", "A>T"="#7570B3", "C>A"="#E7298A", "C>G"="#66A61E", "C>T"="#E6AB02", "del"="#A6761D", "ins"="#666666"))+
    scale_shape_manual(values=c("A>C"=21,"A>G"=21, "A>T"=21, "C>A"=21, "C>G"=21, "C>T"=21, "del"=24, "ins"=25))+
    scale_color_manual(values=c('<5'='lightgrey','5'='black'))+
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          axis.title.x=element_blank(),
          axis.text.x=element_blank(),
          axis.ticks.x=element_blank())+
    geom_segment(mapping = aes(x = starts,xend=starts,y=0,yend=1),data=chrsz,inherit.aes = F,alpha=1/5)+
    geom_text(aes(x=starts+0.5*length,y=0,label=label),data = chrsz,inherit.aes = F)


  # Strelka plot
  cat('Strelka: ',sample)
  temp=unique(strelka_table[sample][,.(cumstart,reads,AFreq,type,SYMBOL,Protein_position,Amino_acids,IMPACT)])
  g_strelka=g_basic+
    xlab(paste('Strelka: ',sample))+
    geom_point(data = temp,mapping = aes(x=cumstart,y=AFreq,fill=type,shape=type,col=reads),alpha=1)+
    geom_label_repel(data = temp[SYMBOL %in% alltier1 & IMPACT %in% c('MODERATE','HIGH'),],
                     mapping = aes(x=cumstart,y=AFreq,label=paste(SYMBOL,Protein_position,Amino_acids)),
                     inherit.aes = F,size=3,nudge_y = 0.2,col='black')+
    geom_label_repel(data = temp[SYMBOL %in% alltier2 & IMPACT %in% c('MODERATE','HIGH'),],
                     mapping = aes(x=cumstart,y=AFreq,label=paste(SYMBOL,Protein_position,Amino_acids)),
                     inherit.aes = F,size=3,nudge_y = 0.2,col='grey')
  print(g_strelka)
}
```


<!-- #### Venn diagram -->
<!-- <!-- Make a Venn diagram of Strelka and Mutect2 findings --> -->
<!-- ```{r mutect2_strelka_venn_diagram, echo=FALSE,,fig.width=6,fig.height=4,warning=F} -->
<!-- if (exists('strelka_indels') & -->
<!--     exists('strelka_snvs') & -->
<!--     exists('mutect2_table')) { -->

<!--   VENN.LIST=list(mutect2=apply(X = mutect2_table[,c('chr','start','end','ref','alt'),], -->
<!--                                MARGIN = 1,FUN = paste,collapse=' '), -->
<!--                  strelka=apply(X = rbind(strelka_snvs[,c('chr','start','end','ref','alt'),], -->
<!--                                          strelka_indels[,c('chr','start','end','ref','alt'),]), -->
<!--                                MARGIN = 1,FUN = paste,collapse=' ')) -->
<!--   venn.plot <- venn.diagram(VENN.LIST , NULL, fill=c("darkmagenta", "darkblue"),  -->
<!--                             alpha=c(0.2,0.2), cex = 2, cat.fontface=4,  -->
<!--                             category.names=c("Mutect 2", "Strelka")) -->
<!--   grid.draw(venn.plot) -->
<!-- } -->
<!-- ``` -->




## Germline point mutations {.tabset}

### HaplotypeCaller (normal only)
<!-- Read annotated small variant data (Haplotype caller, both normal and tumor) -->
```{r load_haplotypecaller, echo=FALSE}
if (exists('haplotypecaller_N_file')) {
  haplotypecaller_files=haplotypecaller_N_file
  if (exists('haplotypecaller_T_file')) haplotypecaller_files=c(haplotypecaller_N_file,haplotypecaller_T_file)
  haplotypecaller_table=NULL
  haplotypecaller_ids=NULL
  for (s in 1:length(haplotypecaller_files)) {
    
    vcf=readVcf(file = haplotypecaller_files[s],genome=reference_genome)
    sample=strsplit(basename(haplotypecaller_files[s]),'[.]')[[1]][1]
    haplotypecaller_ids[[sample]]=names(vcf)
    
    if (s>1) # if not the normal, keep only IDs already present (the normal is first)  <--- this is why somatic variants are not shown
      vcf=vcf[names(vcf) %in% haplotypecaller_table$ID]
    
    # pre-filtering by SWAF / TOPMED to avoid extreme number of variants
    if (project=='BTB') {
      swaf=as.data.table(info(vcf)$SWAF)
      topmed=as.data.table(info(vcf)$TOPMED)
      values=data.table(swaf$value,topmed$value,0) # where both are NA, 0 will be kept as AF
      values$max=apply(values, 1, max, na.rm=T)
      keep=unique(swaf$group[which(values$max<0.01)])  # variant IXs with all alleles <1% in SWAF and TOPMED (swaf$group equals topmed$group)
      vcf=vcf[keep]                 
    } else {
      in_ref=snptable[names(vcf)]
      remove_swegen=which(in_ref$value>=10)
      vcf=vcf[-remove_swegen]
    }
    
    
    
    g=geno(vcf)
    inf=(info(vcf))
    rr=rowRanges(vcf)
    
    mutations=as.data.table(ranges(rr))
    mutations$chr <- as.character(seqnames(rr))
    mutations$sample=sample
    mutations=mutations[,.(ID=names,sample,chr,start,end,width)]
    
    mutations$ref=as.character(ref(vcf))
    mutations$alt=as.data.table(alt(vcf))[, .(values = list(value)), by = group][,values]
    mutations$type=paste0(mutations$ref,'>',mutations$alt)
    mutations$type[nchar(mutations$ref)>nchar(mutations$alt)]='del'
    mutations$type[nchar(mutations$ref)<nchar(mutations$alt)]='ins'
    mutations$type[nchar(mutations$type)>3]='other'
    mutations$type[mutations$type=='T>G']='A>C'
    mutations$type[mutations$type=='T>C']='A>G'
    mutations$type[mutations$type=='T>A']='A>T'
    mutations$type[mutations$type=='G>T']='C>A'
    mutations$type[mutations$type=='G>C']='C>G'
    mutations$type[mutations$type=='G>A']='C>T'
    
    mutations$AFreq=round(sapply(g$AD, "[[", 2) / unlist(g$DP),2) # not a perfect estimate...
    ad=as.data.table(g$AD)
    colnames(ad)='AD'
    mutations=cbind(mutations,ad)
    # mutations$reads=''
    # temp=sapply(g$AD, "[[", 2)
    # mutations$reads[temp<5]='<5'
    # mutations$reads[temp>=5]='5'
    #
    # add counts from the new sources
    if (project=='BTB') suppressWarnings( {
      t=unlist(lapply(X=inf$CAF,FUN=max,na.rm=T))
      t[is.infinite(t)]=0
      mutations$CAF=t
      t=unlist(lapply(X=inf$SWAF,FUN=max,na.rm=T))
      t[is.infinite(t)]=0
      mutations$SWAF=t
      t=unlist(lapply(X=inf$TOPMED,FUN=max,na.rm=T))
      t[is.infinite(t)]=0
      mutations$TOPMED=t
    } )
    
    # "info"" annotations also need to be added
    mutations=cbind(mutations,as.data.table(info(vcf))[,.(CSQ)])
    
    # Add Swegen counts
    mutations$Swegen_count=snptable[names(vcf),value]
    mutations$Swegen_count[is.na(mutations$Swegen_count)]=0
    # Some have multiple rsIDs
    ix=grep(';',mutations$ID)
    ids=data.table(ix,id=strsplit(mutations$ID[ix],';'))
    temp_snptable=snptable[name %in% unlist(ids$id)]
    ids$counts=unlist(lapply(ids$id,function(x) return(max(temp_snptable[x,value],na.rm=T))))
    ids$counts[is.infinite(ids$counts)]=0
    mutations$Swegen_count[ix]=ids$counts
    
    # Also double check Swegen in case pos in reference but rsID or not properly matched in data
    pos=paste0(mutations$chr,':',mutations$start,'_',mutations$ref,'/',mutations$alt)
    snps=snptable[c(pos)][!is.na(value)] # extract from reference[those present]
    if (nrow(snps)>0) mutations$Swegen_count[match(snps$name,pos)]=snps$value # put
    # # Some have multiple alt alleles
    # ix=grep(',',mutations$alt)
    # alts=data.table(ix,alt=mutations$alt[ix])
    # alts$counts[ix]=unlist(lapply(alts$pos,function(x) return(max(snptable[x,value],na.rm=T))))
    
    # More thorough filtering
    if (any(mutations$Swegen_count>=10)) mutations=mutations[-which(Swegen_count>=10)]
    
    
    ## Annotate by cosmic (for possibly retaining non PASS hotspots, which is not necessarily smart)
    key=paste0(substr(mutations$chr,4,6),':',mutations$start,'-',mutations$end)
    key=str_replace(key,'X:','23:')
    key=str_replace(key,'Y:','24:')
    
    if (project!='BTB') key=paste0(mutations$chr,':',mutations$start,'-',mutations$end)
    counts=cbind(cosmic_coding[key,value],cosmic_noncoding[key,value],0)
    max_=apply(counts,1,max,na.rm=T)
    mutations$Cosmic_count=max_
    
    mutations$rank_score=0
    mutations$rank_terms=''
    mutations$LOH=''
    
    # Add Control Freec LOH.
    try( {
      if (nrow(mutations)>0 & !is.null(freec_loh)) for (i in 1:nrow(freec_loh)) {
        ix=freec_loh$chr[i]==mutations$chr &
          freec_loh$start[i]<mutations$start &
          freec_loh$end[i]>mutations$end
        ix=ix[!is.na(ix)]
        if (sum(ix) >0) mutations$LOH[ix]='Y'
      }},silent=T)
    
    
    vep_header=strsplit(info(header(vcf))['CSQ',][,3],'Format: ')[[1]][2]
    vep_header=strsplit(vep_header,'\\|')[[1]]
    
    # VEP annotation is put in annotation_table
    annotation_table=matrix(data = NA,nrow = length(unlist(mutations$CSQ)),ncol = length(vep_header)+1)
    colnames(annotation_table)=c('ID',vep_header)
    row=1
    for (i in 1:nrow(mutations)) { # for each variant
      # for each VEP annotation:
      for (j in 1:length(mutations$CSQ[[i]])) {
        line=strsplit(mutations$CSQ[[i]][j],'\\|')[[1]]
        annotation_table[row,1]=mutations$ID[i]
        annotation_table[row,1+(1:length(line))]=line
        row=row+1
      }
    }
    annotation_table=as.data.table(annotation_table)
    
    # Annotation table then concatenated
    haplotypecaller_table=rbind(haplotypecaller_table,merge(mutations,annotation_table,by='ID'))
  } # done collecting from vcf files
  
  haplotypecaller_table$cumstart=haplotypecaller_table$start
  haplotypecaller_table$cumend=haplotypecaller_table$end
  # # for each chr get cumulative pos
  for (i in 1:nrow(chrsz)) {
    ix=haplotypecaller_table$chr==chrsz$chr[i]
    haplotypecaller_table$cumstart[ix]=haplotypecaller_table$start[ix]+chrsz$starts[i]
    haplotypecaller_table$cumend[ix]=haplotypecaller_table$end[ix]+chrsz$starts[i]
  }
  setkey(haplotypecaller_table,'sample')
  
  # Due to size: remove intron/intergenic variants
  selection=haplotypecaller_table[!Consequence %in% c('intron_variant','intergenic_variant'),-c('CSQ')] # not needed after parsing/merging the annotations
  
  if (nrow(selection)>0) {
    
    # Cosmic/local Tier2 :
    ix=selection$SYMBOL %in% alltier2
    selection$rank_score[ix]=2
    selection$rank_terms[ix]='T2_gene'
    # tier 1 priority:
    ix=selection$SYMBOL %in% alltier1
    selection$rank_score[ix]=2
    selection$rank_terms[ix]='T1_gene'
    
    # Add high impact
    ix=selection$IMPACT=='HIGH'
    if (any(ix)) {
      selection$rank_score[ix]=selection$rank_score[ix]+2
      selection$rank_terms[ix]=paste(selection$rank_terms[ix],'high_impact')
    }
    
    # Add moderate impact
    ix=selection$IMPACT=='MODERATE'
    if (any(ix)) {
      selection$rank_score[ix]=selection$rank_score[ix]+1
      selection$rank_terms[ix]=paste(selection$rank_terms[ix],'moderate_impact')
    }
    
    # additional +1 if high impact and TSG
    ix=selection$IMPACT=='HIGH' & selection$SYMBOL %in% alltsg
    if (any(ix)) {
      selection$rank_score[ix]=selection$rank_score[ix]+1
      selection$rank_terms[ix]=paste(selection$rank_terms[ix],'high+TSG')
    }
    
    # Add clinvar pathogenic
    ix=grep('pathogenic',selection$CLIN_SIG)
    if (length(ix)>0) {
      selection$rank_score[ix]=selection$rank_score[ix]+2
      selection$rank_terms[ix]=paste(selection$rank_terms[ix],'clinvar')
    }
    
    # Add Polyphen/SIFT damaging/deleterious
    ix=union(grep('damaging',selection$PolyPhen),grep('deleterious',selection$SIFT))
    if (length(ix)>0) {
      selection$rank_score[ix]=selection$rank_score[ix]+1
      selection$rank_terms[ix]=paste(selection$rank_terms[ix],'polyphen/SIFT')
    }
    
    # Add hotspots and near hotspots (within 2 residues of a hotspot)
    key=paste(selection$SYMBOL,
              str_replace(string = selection$Protein_position,
                          pattern = '/.*',replacement = ''))
    ix <-
      key %in% paste(hotspots_snv[,Hugo_Symbol],hotspots_snv[,Amino_Acid_Position]) |
      key %in% paste(hotspots_inframe[,Hugo_Symbol],hotspots_inframe[,Amino_Acid_Position])  ## Warning: Exact match used with inframes
    if (any(ix)) {
      selection$rank_score[ix]=selection$rank_score[ix]+2
      selection$rank_terms[ix]=paste(selection$rank_terms[ix],'hotspot')
    }
    ix = !ix & key %in% near_hotspots # the near_hotspot excludes those that were a hotspot
    if (any(ix)) {
      selection$rank_score[ix]=selection$rank_score[ix]+1
      selection$rank_terms[ix]=paste(selection$rank_terms[ix],'near_hotspot')
    }
    
    # Add cosmic counts
    ix=which(selection$Cosmic_count>50)
    if (length(ix)>0) {
      selection$rank_score[ix]=selection$rank_score[ix]+2
      selection$rank_terms[ix]=paste(selection$rank_terms[ix],'cosmic_>50')
    }
    ix=which(selection$Cosmic_count>5 & selection$Cosmic_count<=50)
    if (length(ix)>0) {
      selection$rank_score[ix]=selection$rank_score[ix]+1
      selection$rank_terms[ix]=paste(selection$rank_terms[ix],'cosmic_>5')
    }
    
    # Special case for TERT promoter
    ix=which(selection$SYMBOL=='TERT' & selection$Consequence=='5_prime_UTR_variant')
    if (length(ix)>0) {
      selection$rank_score[ix]=selection$rank_score[ix]+2
      selection$rank_terms[ix]=paste(selection$rank_terms[ix],'TERT_5\'UTR')
    }

    # Add TF binding variants near (100kb) cancer genes
    ix=grep('TF',selection$Consequence)
    if (length(ix)>0) {
      selection$rank_score[ix]=selection$rank_score[ix]+2
      selection$rank_terms[ix]=paste(selection$rank_terms[ix],'TFBS')
      for (i in 1:length(ix)) {
        near=which(selection$cumend[ix[i]] > (tumorgenes$cumstart-100e3) & selection$cumstart[ix[i]] < (tumorgenes$cumend + 100e3) &
                     !selection$SYMBOL[ix[i]] %in% alltumorgenes)
        if (length(near)>0) {
          genes=paste(tumorgenes$`Gene Symbol`[near],collapse = ',')
          selection$rank_score[ix[i]]=selection$rank_score[ix[i]]+2
          selection$rank_terms[ix[i]]=paste(selection$rank_terms[ix[i]],paste0('near_',genes))
        }
      }
    }
    
    ix=which(selection$CANONICAL!='YES')
    if (length(ix)>0) {
      selection$rank_score[ix]=selection$rank_score[ix]-0
      selection$rank_terms[ix]=paste(selection$rank_terms[ix],'not_canonical')
    }
  }
  
  firstcols=c('ID','sample','SYMBOL','rank_score','rank_terms','LOH','AFreq','Consequence','IMPACT','SWAF','TOPMED','Swegen_count')
  cols=colnames(selection)
  setcolorder(x = selection,neworder = c(firstcols,cols[!cols %in% firstcols]))
  
  haplotypecaller_selected <- selection[order(cumstart,Allele)][order(rank_score,decreasing = T)]
  if (write_tables) fwrite(haplotypecaller_selected,file=paste0(sampleData$name,'_haplotypecaller.csv'))
  ix=haplotypecaller_selected$sample==strsplit(basename(haplotypecaller_files[1]),'[.]')[[1]][1] # first file is the normal
  tableWrapper(haplotypecaller_selected[ix][,-c('cumstart','cumend','DOMAINS')][rank_score>1])
  
}

```

### HaplotypeCaller (all samples)
```{r haplotypecaller_T+N, echo=FALSE}
if (project=='BTB') tableWrapper(haplotypecaller_selected[,-c('cumstart','cumend','DOMAINS')][rank_score>1])
```

### HaplotypeCaller overlap
```{r haplotypecaller_overlap, echo=FALSE}
# Show the overlap
if (exists('haplotypecaller_N_file')) {
  names=names(haplotypecaller_ids)
  nsamples=length(names)
  table=matrix('',ncol=nsamples,nrow=nsamples,
               dimnames=list(`Variants in`=names,
                             `Also in`=names))
  for (i in 1:nsamples) for (j in 1:nsamples) {
    s=sum(haplotypecaller_ids[[names[i]]] %in% haplotypecaller_ids[[names[j]]])
    table[i,j]=paste(round(s/1e6,2),'M')
  }
  htmlTable(table)
}
```



## All {.tabset}

### Top ranked {.active}
```{r}
# make several small tables, from which results can be copy-pasted.
# each is just a subset (incl sample id) if the sorted "selected" variants.
if (exists('mutect2_selected')) {
  cat('Mutect2')
  t=mutect2_selected[rank_score>0]
  if (nrow(t)>0){
    t=data.table(Sample=t$sample,
                 Gene=t$SYMBOL,
                 Mutation=paste(t$Protein_position,t$Amino_acids),
                 'Rank score'=t$rank_score,
                 'Rank Terms'=t$rank_terms,
                 Allele_ratio=t$AFreq)
    kable(t[Gene!=''], "html") %>%
      kable_styling(bootstrap_options = "striped",position = "left") %>%
      scroll_box(width = "900px", height = "300px")

  }}
```

&nbsp;

```{r}
cat('Strelka')
t=strelka_selected[rank_score>0]
if (nrow(t)>0){
  t=data.table(Sample=t$sample,
               Gene=t$SYMBOL,
               Mutation=paste(t$Protein_position,t$Amino_acids),
               'Rank score'=t$rank_score,
               'Rank Terms'=t$rank_terms,
               Allele_ratio=t$AFreq)
  kable(t[Gene!=''], "html") %>%
  kable_styling(bootstrap_options = "striped",position = "left") %>%
  scroll_box(width = "900px", height = "300px")
}
```

&nbsp;

```{r}

if (exists('haplotypecaller_selected')) {
  cat('HaplotypeCaller')
  t=haplotypecaller_selected[rank_score>1]
  if (nrow(t)>0){
    t=data.table(Sample=t$sample,
                 Gene=t$SYMBOL,
                 Mutation=paste(t$Protein_position,t$Amino_acids),
                 'Rank score'=t$rank_score,
                 'Rank Terms'=t$rank_terms,
                 Allele_ratio=t$AFreq)
    kable(t[Gene!=''], "html") %>%
      kable_styling(bootstrap_options = "striped",position = "left") %>%
      scroll_box(width = "900px", height = "300px")
  }
}
```

&nbsp;

```{r}
cat('Manta somatic')
try( {
  t=manta_tumor_selected[rank_score>0]
  t$Gene_Name[nchar(t$Gene_Name)>100]='many'
  if (nrow(t)>0){
    t=unique(data.table(Sample=t$sample,
                 Gene=t$Gene_Name,
                 Mutation=t$SVTYPE,
                 'Rank score'=t$rank_score,
                 'Rank Terms'=t$rank_terms,
                 Allele_ratio=t$AFreq))
    kable(t[Gene!=''], "html") %>%
      kable_styling(bootstrap_options = "striped",position = "left") %>%
      scroll_box(width = "900px", height = "300px")
  }
}, silent=T)
```

&nbsp;

```{r}
cat('Manta germline')
if (exists('manta_normal_selected')) try({
  t=manta_normal_selected[rank_score>1]
  t$Gene_Name[nchar(t$Gene_Name)>100]='many'
  if (nrow(t)>0){
    t=unique(data.table(Sample=t$sample,
                 Gene=t$Gene_Name,
                 Mutation=t$SVTYPE,
                 'Rank score'=t$rank_score,
                 'Rank Terms'=t$rank_terms,
                 Allele_ratio=t$AFreq))
    kable(t[Gene!=''], "html") %>%
      kable_styling(bootstrap_options = "striped",position = "left") %>%
      scroll_box(width = "900px", height = "300px")
  }
},silent=T)
```

&nbsp;

```{r}

if (exists('freec_cnv')) try({
  t=freec_cnv
  if (!is.null(t)) if (nrow(t)>0){
    cat('Control Freec')
    t=data.table(Sample=t$sample,
                 Gene=t$cancer_genes,
                 Mutation=t$effect,
                 'Rank score'=t$rank_score,
                 'Rank Terms'=t$rank_terms,
                 Allele_ratio='')
    kable(t[Gene!=''], "html") %>%
      kable_styling(bootstrap_options = "striped",position = "left") %>%
      scroll_box(width = "900px", height = "300px")
  }
}, silent=T)
```

&nbsp;

```{r}
cat('Ascat')
try( {
  t=ascat_cnv
  if (nrow(t)>0){
    t=data.table(Sample=t$sample,
                 Gene=t$cancer_genes,
                 Mutation=paste(t$nMajor,t$nMinor),
                 'Rank score'=t$rank_score,
                 'Rank Terms'=t$rank_terms,
                 Allele_ratio='')
    kable(t[Gene!=''], "html") %>%
      kable_styling(bootstrap_options = "striped",position = "left") %>%
      scroll_box(width = "900px", height = "300px")
  }}, silent=T)
```

&nbsp;

### Save to files
```{r save_to_files}
#export bed file of all relevant genomic regions
if (T) if (write_tables) try({
  x = unique(rbind(
      manta_tumor_table[,.(chr,start,end=start)],
      manta_tumor_table[,.(chr,start=end,end)],
      manta_tumor_table[!is.na(altpos),.(chr=altchr,start=altpos,end=altpos)],
      manta_normal_table[,.(chr,start,end=start)],
      manta_normal_table[,.(chr,start=end,end)],
      manta_normal_table[!is.na(altpos),.(chr=altchr,start=altpos,end=altpos)],
      mutect2_table[,.(chr,start,end)],
      strelka_table[,.(chr,start,end)],
      haplotypecaller_table[,.(chr,start,end)])
    )
  #x[,1]=paste0('chr',x[,1])
  #save(x,file =  paste0(sampleData$name,'_forBamfile.Rdata'))
  write.table(
    x,
    file = paste0(sampleData$name,'_forBamfile.bed'
    ),
    sep = '\t',row.names = F,col.names = F,quote=F)
}, silent=T)

save.image(paste0(sampleData$name,'_ws.Rdata'))

```



## Chromosomes {.tabset}
<!-- Plot composite data, chromosome wise (updated but does not work yet, the old function below is still used) -->
```{r old_plot_function}
chromplot=function(thischr) if (T) {
  if (project=='BTB') thischr=paste0('chr',thischr)
  samplenames=sort(unique(ascat_binned$sample))
  nsamples=length(samplenames)
  for (s in 1:nsamples) {
    cat(samplenames[s])
    par(mai=c(0,4,0,2))
    g=ggplot()+
      scale_y_continuous(limits=c(-2.7,3),
                         breaks = seq(1.5,-2.5,-0.5),
                         minor_breaks = seq(1.25,-2.25,-0.5),
                         labels =
                           c('Tumor DNA=1.5',
                             'Tumor Copy Ratio\n(median centered)',
                             'Tumor DNA=0.5\nMutation AF=1',
                             'Mutation AF=0.5',
                             'Mutation AF=0\nTumor SNP AF=1',
                             'Tumor\nSNPs',
                             'Tumor SNP AF=0\nNormal SNP AF=1',
                             'Normal\nSNPs',
                             'Normal SNP AF=0'))+
      scale_x_continuous(
        breaks = 10e6*seq(0,ceiling(chrsz$length[chrsz$chr==thischr]/10e6)),
        labels= c('0',paste0(10*seq(1,ceiling(chrsz$length[chrsz$chr==thischr]/10e6)),'M')),
        minor_breaks = 1e6*seq(0,ceiling(chrsz$length[chrsz$chr==thischr]/1e6)))+
      theme(panel.grid.minor.x = element_line(colour="lightgrey", size=0.05),
            panel.grid.major.x = element_line(colour="grey", size=0.2),
            panel.grid.major.y = element_line(colour="black", size=0.2),
            panel.grid.minor.y = element_line(colour="lightgrey", size=0.05),
            axis.title.x=element_blank(),
            axis.ticks.x=element_blank(),
            axis.title.y=element_blank(),
            axis.ticks.y=element_blank())#+

    # lines to separate parts of fig
    g=g+geom_segment(mapping = aes(x = 0,xend=chrsz$length[chrsz$chr==thischr],y=0.5,yend=0.5),alpha=1)
    g=g+geom_segment(mapping = aes(x = 0,xend=chrsz$length[chrsz$chr==thischr],y=-0.5,yend=-0.5),alpha=1)
    g=g+geom_segment(mapping = aes(x = 0,xend=chrsz$length[chrsz$chr==thischr],y=-1.5,yend=-1.5),alpha=1)
    g=g+geom_segment(mapping = aes(x = 0,xend=chrsz$length[chrsz$chr==thischr],y=-2.5,yend=-2.5),alpha=1)
    #g=g+geom_rect(mapping = aes(ymin=-0.5,ymax=0.5,xmin=0,xmax=1e6*ceiling(chrsz$length[chrsz$chr==thischr]/1e6)),alpha=1,col=NA)

    ### TUMOR (& normal) LOGRATIO
    if (exists('tratio')) {
      temp <- nratio
      temp$Ratio[temp$Ratio>3]=3
      g=g+geom_line(aes(x=Start,y=Ratio),
                    data = temp[temp$Chromosome==thischr,],alpha=1/3,
                    stat="identity",colour="red")
      samples=sort(unique(tratio$sample))
      temp <- tratio[samples[s]]
      temp$Ratio[temp$Ratio>3]=3
      g=g+geom_line(aes(x=Start,y=Ratio),
                    data = temp[temp$Chromosome==thischr,],alpha=2/3,
                    stat="identity",colour="black")
    } else if (exists('ascat_tratio')) {
      samples=sort(unique(ascat_tratio$sample))
      temp <- ascat_tratio[samples[s]]
      temp$smoothed[temp$smoothed>3]=3
      g=g+geom_line(aes(x=Position,y=smoothed),
                    data = temp[temp$Chr==thischr,],alpha=1/3,
                    stat="identity",colour="black")
    }

    ### TUMOR STRVAR
    if (!is.null(manta_tumor_table)) if (nrow(manta_tumor_table)>0) {
      rm(chr)
      samples=sort(unique(manta_tumor_table$sample))
      # first same chrom SVs
      temp=unique(manta_tumor_table[sample==samples[s] & chr==thischr & chr==altchr,
                                       .(sample,chr,altchr,start,altpos,IMPRECISE,SVTYPE)])
      if (nrow(temp)>0) g=g+geom_curve(aes(x = start,xend = altpos,y=1.99,yend=2,col=SVTYPE,linetype=IMPRECISE),
                                       data = temp,alpha=1/2,
                                       curvature=-0.2,size=1,
                                       position = "identity", angle = 90, ncp = 50,
                                       arrow = arrow(length=unit(0.10,"cm"), ends="both", type = "open"),
                                       lineend = "butt", na.rm = FALSE, show.legend = NA,
                                       inherit.aes = FALSE)+
          scale_color_manual(values=c("BND"="#E41A1C", "DEL"="#377EB8", "INV"="#4DAF4A", "INS"="#984EA3", "DUP"="#FF7F00"))+
          scale_linetype_manual(values=c('TRUE'=2,'FALSE'=1))
      
      # then diff chrom SVs with other end towards left
      temp=unique(manta_tumor_table[sample==samples[s] & chr==thischr & altcumpos<cumstart & chr!=altchr,
                                    .(sample,chr,altchr,start,IMPRECISE,SVTYPE)])
      if (nrow(temp)>0) {
        g=g+geom_curve(aes(x = 0,xend = start,y=3,yend=2,col=SVTYPE,linetype=IMPRECISE),
                       data = temp,alpha=1/2,
                       curvature=-0.2,size=1,
                       position = "identity", angle = 90, ncp = 50,
                       arrow = arrow(length=unit(0.10,"cm"), ends="both", type = "open"),
                       lineend = "butt", na.rm = FALSE, show.legend = NA,
                       inherit.aes = FALSE)
      }
      # then diff chrom SVs with other end towards right
      temp=unique(manta_tumor_table[sample==samples[s] & chr==thischr & altcumpos>cumstart & chr!=altchr,
                                    .(sample,chr,altchr,start,IMPRECISE,SVTYPE)])
      
      if (nrow(temp)>0) {
        g=g+geom_curve(aes(x = start,xend = chrsz$length[chrsz$chr==thischr],y=2,yend=3,col=SVTYPE,linetype=IMPRECISE),
                       data = temp,alpha=1/2,
                       curvature=-0.2,size=1,
                       position = "identity", angle = 90, ncp = 50,
                       arrow = arrow(length=unit(0.10,"cm"), ends="both", type = "open"),
                       lineend = "butt", na.rm = FALSE, show.legend = NA,
                       inherit.aes = FALSE)
      }
    }
      
      
    ### TUMOR MUTATIONS
    samples=sort(unique(strelka_table$sample))
    temp=unique(strelka_table[samples[s]][chr==thischr & type!='other',][,.(start,AFreq,type,Amino_acids,SYMBOL,IMPACT,Protein_position)])
    if (nrow(temp)>0) {
      g=g+geom_point(aes(x=start,y=AFreq -0.5,fill=type,shape=type),data = temp,alpha=1)+
        scale_fill_manual(values = c("A>C"="#1B9E77","A>G"="#D95F02", "A>T"="#7570B3", "C>A"="#E7298A", "C>G"="#66A61E", "C>T"="#E6AB02", "del"="#A6761D", "ins"="#666666"))+
        scale_shape_manual(values=c("A>C"=21,"A>G"=21, "A>T"=21, "C>A"=21, "C>G"=21, "C>T"=21, "del"=24, "ins"=25))#+

      # with labels
      g=g+geom_label_repel(data = temp[IMPACT %in% c('MODERATE','HIGH') & SYMBOL %in% alltier2,],
                           mapping = aes(x=start,y=AFreq -0.5,
                                         label=paste(SYMBOL,Protein_position,Amino_acids)),
                           inherit.aes = F,size=3,min.segment.length = 0,nudge_y = 0.3,col='grey')
      g=g+geom_label_repel(data = temp[IMPACT %in% c('MODERATE','HIGH') & SYMBOL %in% alltier1,],
                           mapping = aes(x=start,y=AFreq -0.5,
                                         label=paste(SYMBOL,Protein_position,Amino_acids)),
                           inherit.aes = F,size=3,min.segment.length = 0,nudge_y = 0.3,col='black')
    }

    ### SNPS
    if (exists('tbaf')) {
      samples=sort(unique(tbaf$sample))
      g=g+geom_point(aes(x=Position,y=BAF -1.5),data = tbaf[sample==samples[s]][Chromosome==thischr],alpha=1/8,shape='.')
    } else if (exists('ascat_tbaf'))  {
      samples=sort(unique(ascat_tbaf$sample))
      g=g+geom_point(aes(x=Position,y=BAF -1.5),data = ascat_tbaf[sample==samples[s]][Chr==thischr],alpha=1/8,shape='.')
    }
    if (exists('nbaf')) {
      g=g+geom_point(aes(x=Position,y=BAF -2.5),data = nbaf[nbaf$Chromosome==thischr,],alpha=1/8,shape='.')
    } else if (exists('ascat_nbaf')) {
      g=g+geom_point(aes(x=Position,y=BAF -2.5),data = ascat_nbaf[ascat_nbaf$Chr==thischr,],alpha=1/8,shape='.')
    }
    ## Germline muts
    if (exists('haplotypecaller_selected')) {
      if (exists('haplotypecaller_T_file')) {
        sample=strsplit(basename(haplotypecaller_T_file[s]),'[.]')[[1]][1]
        temp=haplotypecaller_selected[sample==sample][chr==thischr][rank_score>1][type!='other']
        temp=unique(temp[,.(start,type,AFreq,Amino_acids,Protein_position,SYMBOL)])
        g=g+
        geom_point(aes(x=start,y=AFreq -1.5,fill=type,shape=type),
                   data = temp,alpha=1)+
        geom_label_repel(data = temp[Amino_acids!='' & SYMBOL %in% alltumorgenes],
                         mapping = aes(x=start,y=AFreq -1.5,label=paste(SYMBOL,Protein_position,Amino_acids)),
                         inherit.aes = F,size=3,min.segment.length = 0)
      }
      normal=strsplit(basename(haplotypecaller_N_file),'[.]')[[1]][1]
      temp=haplotypecaller_selected[sample==normal][chr==thischr][rank_score>1][type!='other']
      temp=unique(temp[,.(start,type,AFreq,Amino_acids,Protein_position,SYMBOL)])
      g=g+
        geom_point(aes(x=start,y=AFreq -2.5,fill=type,shape=type),
                   data = temp,alpha=1)+
        geom_label_repel(data = temp[Amino_acids!='' & SYMBOL %in% alltumorgenes],
                         mapping = aes(x=start,y=AFreq -2.5,label=paste(SYMBOL,Protein_position,Amino_acids)),
                         inherit.aes = F,size=3,min.segment.length = 0)
    }

    #genes!
    temp=tumorgenes[chr==thischr]
    if (nrow(temp)>0) g=g+geom_text_repel(mapping = aes(x = (start+end)/2,y =  -2.6,label=`Gene Symbol`),
                        data=temp,
                        nudge_y= -0.1,size=2)

    print(g)
  }
}
```

```{r chromosomes_plot_function}
chromplot2=function(thischr) {
  # basic plot element:
  g0=ggplot()+
    scale_x_continuous(
      breaks = 10e6*seq(0,ceiling(chrsz$length[chrsz$chr==thischr]/10e6)),
      labels= c('0',paste0(10*seq(1,ceiling(chrsz$length[chrsz$chr==thischr]/10e6)),'M')),
      minor_breaks = 1e6*seq(0,ceiling(chrsz$length[chrsz$chr==thischr]/1e6))
    )+
    theme(
      panel.grid.minor.x = element_line(colour="lightgrey", size=0.05),
      panel.grid.major.x = element_line(colour="grey", size=0.2),
      panel.grid.major.y = element_line(colour="black", size=0.2),
      panel.grid.minor.y = element_line(colour="lightgrey", size=0.05),
      axis.title.x=element_blank(),
      axis.ticks.x=element_blank(),
      axis.title.y=element_blank(),
      axis.ticks.y=element_blank()
    )


  #copy number element
  g_cnv=NULL
  g=g0+ylim(y=c(0,3))
  if (exists('tratio')) for (sample in sort(unique(tratio$sample))) {
    cat('Control Freec,',sample)
    temp <- nratio
    temp$Ratio[temp$Ratio>3]=3
    g=g+geom_line(aes(x=Start,y=Ratio),
                  data = temp[temp$Chromosome==thischr,],alpha=1/3,
                  stat="identity",colour="red")
    temp <- tratio
    temp$Ratio[temp$Ratio>3]=3
    g=g+geom_line(aes(x=Start,y=Ratio),
                  data = temp[temp$Chromosome==thischr,],alpha=2/3,
                  stat="identity",colour="black")
    g_cnv[[sample]]=g
  } else if (exists('ascat_tratio')) for (sample in sort(unique(ascat_tratio$sample))) {
    cat('Ascat,', sample)
    temp <- ascat_tratio[sample]
    temp$smoothed[temp$smoothed>3]=3
    g=g+geom_line(aes(x=Position,y=smoothed),
                  data = temp[temp$Chr==thischr,],alpha=1/3,
                  stat="identity",colour="black")
    g_cnv[[sample]]=g
  }


  # structutal variants element
  g_sv=NULL
  g=g0+ylim(c(-1,1))
  # if (exists('tratio')) {
  #   temp <- tratio
  #   temp$Ratio[temp$Ratio>3]=3
  #   g=g+geom_line(aes(x=Start,y=log2(Ratio)),
  #                 data = temp[temp$Chromosome==thischr,],alpha=1/10,
  #                 stat="identity",colour="black")
  # } else if (exists('ascat_tratio')) {
  #   temp <- ascat_tratio
  #   temp$smoothed[temp$smoothed>3]=3
  #   g=g+geom_line(aes(x=Position,y=log2(smoothed)),
  #                 data = temp[temp$Chr==thischr,],alpha=1/10,
  #                 stat="identity",colour="black")
  # }
  if (exists('manta_tumor_selected')) if (nrow(manta_tumor_selected)>0) for (sample in sort(unique(manta_tumor_selected$sample))) {
    cat('Manta:',sample)
    setkey(manta_tumor_selected,'sample')
    sv_table=unique(manta_tumor_selected[sample][chr==thischr,.(sample,chr,altchr,cumstart,altcumpos,IMPRECISE,SVTYPE,plot)])

    if (nrow(sv_table)>0) {
    # first same chrom SVs
    temp=sv_table[altchr==chr,]
    if (nrow(temp)>0) g=g+geom_curve(aes(x = start,xend = altpos,y=0,yend=0.001,col=SVTYPE,linetype=IMPRECISE),
                                     data = temp,alpha=1/2,
                                     curvature=-0.2,size=1,
                                     position = "identity", angle = 90, ncp = 50,
                                     arrow = arrow(length=unit(0.10,"cm"), ends="both", type = "open"),
                                     lineend = "butt", na.rm = FALSE, show.legend = NA,
                                     inherit.aes = FALSE)+
        scale_color_manual(values=c("BND"="#E41A1C", "DEL"="#377EB8", "INV"="#4DAF4A", "INS"="#984EA3", "DUP"="#FF7F00"))+
        scale_linetype_manual(values=c('TRUE'=2,'FALSE'=1))

  }
  g_sv[[sample]]=g

  # somatic small variants element
  g=g0+ylim(y=c(0,1))
  if (exists('mutect2_table')) if (nrow(mutect2_table)>0) {
    temp=mutect2_table[chr==thischr & filter=='PASS' & Callers=='Mutect 2 and Strelka',]
    if (nrow(temp)>0) {
      g=g+geom_point(aes(x=start,y=AFreq,fill=type,shape=type),data = temp,alpha=1)+
        scale_fill_manual(
          values=c("A>C"="#1B9E77","A>G"="#D95F02", "A>T"="#7570B3",
                   "C>A"="#E7298A", "C>G"="#66A61E", "C>T"="#E6AB02",
                   "del"="#A6761D", "ins"="#666666"))+
        scale_shape_manual(
          values=c("A>C"=21,"A>G"=21, "A>T"=21,
                   "C>A"=21, "C>G"=21, "C>T"=21,
                   "del"=24, "ins"=25))+
        scale_color_manual(values=c('<5'='lightgrey','5'='black'))

      g=g+geom_label_repel(
        data = temp[temp$Amino_acids!='',],
        mapping = aes(x=start,y=AFreq,label=paste(SYMBOL,Protein_position,Amino_acids)),
        inherit.aes = F,size=3,min.segment.length = 0,nudge_y = 0
      )
    }
  }
  g_mut=g

  # tumor SNPs
  g=g0+ylim(y=c(0,1))
  if (exists('tbaf')) {
    g=g+geom_point(aes(x=Position,y=BAF),data = tbaf[tbaf$Chromosome==thischr,],alpha=1/8,shape='.')
  } else if (exists('ascat_tbaf'))  {
    g=g+geom_point(aes(x=Position,y=BAF),data = ascat_tbaf[ascat_tbaf$Chr==thischr,],alpha=1/8,shape='.')
  }
  if (exists('haplotypecaller_table')) {
      temp=haplotypecaller_table[keep_haplotypecaller][chr==thischr]
      g=g+
        geom_point(aes(x=start,y=AF_t,fill=type,shape=type),
                   data = temp,alpha=1)+
        geom_label_repel(data = temp[Amino_acids!='',],
                         mapping = aes(x=start,y=AF_t,label=paste(SYMBOL,Protein_position,Amino_acids)),
                         inherit.aes = F,size=3,min.segment.length = 0)+
        scale_fill_manual(
          values=c("A>C"="#1B9E77","A>G"="#D95F02", "A>T"="#7570B3",
                   "C>A"="#E7298A", "C>G"="#66A61E", "C>T"="#E6AB02",
                   "del"="#A6761D", "ins"="#666666"))+
        scale_shape_manual(
          values=c("A>C"=21,"A>G"=21, "A>T"=21,
                   "C>A"=21, "C>G"=21, "C>T"=21,
                   "del"=24, "ins"=25))
  }
  g_tsnp=g

  # normal SNPs
  g=g0+ylim(y=c(0,1))
  if (exists('nbaf')) {
    g=g+geom_point(aes(x=Position,y=BAF),data = nbaf[nbaf$Chromosome==thischr,],alpha=1/8,shape='.')
  } else if (exists('ascat_nbaf')) {
    g=g+geom_point(aes(x=Position,y=BAF),data = ascat_nbaf[ascat_nbaf$Chr==thischr,],alpha=1/8,shape='.')
  }
  if (exists('haplotypecaller_table')) {
    g=g+
      geom_point(aes(x=start,y=AF,fill=type,shape=type),
                 data = temp,alpha=1)+
      geom_label_repel(data = temp[Amino_acids!='',],
                       mapping = aes(x=start,y=AF,label=paste(SYMBOL,Protein_position,Amino_acids)),
                       inherit.aes = F,size=3,min.segment.length = 0)+
        scale_fill_manual(
          values=c("A>C"="#1B9E77","A>G"="#D95F02", "A>T"="#7570B3",
                   "C>A"="#E7298A", "C>G"="#66A61E", "C>T"="#E6AB02",
                   "del"="#A6761D", "ins"="#666666"))+
        scale_shape_manual(
          values=c("A>C"=21,"A>G"=21, "A>T"=21,
                   "C>A"=21, "C>G"=21, "C>T"=21,
                   "del"=24, "ins"=25))
  }
  g_nsnp=g

  grid.newpage()
  grid.draw(rbind(ggplotGrob(g_cnv),
                  ggplotGrob(g_sv),
                  ggplotGrob(g_mut),
                  ggplotGrob(g_tsnp),
                  ggplotGrob(g_nsnp),size = "last"))

}} # end of plot function
```


### 1
```{r plot_chr1,fig.width=15,fig.height=10}
chr=1
cat('Chromosome',chr)
chromplot(chr)
```

### 2
```{r plot_chr2,fig.width=15,fig.height=10}
chr=2
cat('Chromosome',chr)
chromplot(chr)
```

### 3
```{r plot_chr3,fig.width=15,fig.height=10}
chr=3
cat('Chromosome',chr)
chromplot(chr)
```

### 4
```{r plot_chr4,fig.width=15,fig.height=10}
chr=4
cat('Chromosome',chr)
chromplot(chr)
```

### 5
```{r plot_chr5,fig.width=15,fig.height=10}
chr=5
cat('Chromosome',chr)
chromplot(chr)
```

### 6
```{r plot_chr6,fig.width=15,fig.height=10}
chr=6
cat('Chromosome',chr)
chromplot(chr)
```

### 7
```{r plot_chr7,fig.width=15,fig.height=10}
chr=7
cat('Chromosome',chr)
chromplot(chr)
```

### 8
```{r plot_chr8,fig.width=15,fig.height=10}
chr=8
cat('Chromosome',chr)
chromplot(chr)
```

### 9
```{r plot_chr9,fig.width=15,fig.height=10}
chr=9
cat('Chromosome',chr)
chromplot(chr)
```

### 10
```{r plot_chr10,fig.width=15,fig.height=10}
chr=10
cat('Chromosome',chr)
chromplot(chr)
```

### 11
```{r plot_chr11,fig.width=15,fig.height=10}
chr=11
cat('Chromosome',chr)
chromplot(chr)
```

### 12
```{r plot_chr12,fig.width=15,fig.height=10}
chr=12
cat('Chromosome',chr)
chromplot(chr)
```

### 13
```{r plot_chr13,fig.width=15,fig.height=10}
chr=13
cat('Chromosome',chr)
chromplot(chr)
```

### 14
```{r plot_chr14,fig.width=15,fig.height=10}
chr=14
cat('Chromosome',chr)
chromplot(chr)
```

### 15
```{r plot_chr15,fig.width=15,fig.height=10}
chr=15
cat('Chromosome',chr)
chromplot(chr)
```

### 16
```{r plot_chr16,fig.width=15,fig.height=10}
chr=16
cat('Chromosome',chr)
chromplot(chr)
```

### 17
```{r plot_chr17,fig.width=15,fig.height=10}
chr=17
cat('Chromosome',chr)
chromplot(chr)
```

### 18
```{r plot_chr18,fig.width=15,fig.height=10}
chr=18
cat('Chromosome',chr)
chromplot(chr)
```

### 19
```{r plot_chr19,fig.width=15,fig.height=10}
chr=19
cat('Chromosome',chr)
chromplot(chr)
```

### 20
```{r plot_chr20,fig.width=15,fig.height=10}
chr=20
cat('Chromosome',chr)
chromplot(chr)
```

### 21
```{r plot_chr21,fig.width=15,fig.height=10}
chr=21
cat('Chromosome',chr)
chromplot(chr)
```

### 22
```{r plot_chr22,fig.width=15,fig.height=10}
chr=22
cat('Chromosome',chr)
chromplot(chr)
```

### X
```{r plot_chrX,fig.width=15,fig.height=10}
chr='X'
cat('Chromosome',chr)
chromplot(chr)
```

### Y
```{r plot_chrY,fig.width=15,fig.height=10}
chr='Y'
cat('Chromosome',chr)
chromplot(chr)
```

